
## Do prices vary by size? {#AppendixHedonic}
<!--- 
The hedonic model is estimated by 
global in_string 2026-03-03
do "$analysis_code/bsb_size_classifications.do"
It creates a table and 3 figures that appear in this appendix
--->
We first estimate a linear hedonic model in which the per-pound price of a lot of fish is a function of the attributes of that lot, including  market category, gear, and time of year. @McConnell2000 and @Carroll2001 are early examples of hedonic models applied to tuna prices: attributes like freshness, fat content, and individual size are associated with positive price premia.  @Kristofersson2004 [@Kristofersson2007], and @Hammarlund2015 examine demand for attributes in hierarchical linear model and find that, for example, large fish receive higher prices but when the quantity supplied of large fish is higher than average, the price premia is smaller than the average. Other attributes can impact prices: See @Gobillon2017 for an examination of buyer- and seller- effects. 

We use data from 2013-2025 aggregated to the "trip, market-category, gear, area, datesail, dateland, dealer" level and deflate to real prices using the CPI-U for the first quarter of 2023.  We perform some aggregation of gears: Line and Hand, Trawl, Gillnet, Pot and Trap and "All other" gears.  We use identical market category aggregations as in the Random Forest.  Table \@ref(tab:bsb-hedonic-summary) summarizes the data used to estimate the hedonic model. 

We include transactions from New Hampshire to North Carolina where the nominal price is between \$0.15 and \$12 per pound (\$0.33 to \$5.44 per kg). The quantity of fish in each transaction varies and the coefficients from the unweighted regression can be interpreted as the marginal effects of changes in the independent variables on the average price of a sales transaction.   We also estimate a model in which each transaction is weighted by pounds landed.  Those coefficients can be interpreted as marginal effects on the average price of a pound of Black Sea Bass.  In addition to changing the interpretation, the weighting carries an extra benefit.  The algorithm that creates CAMS data allocates quantities on from multi-area or multi-gear trips to those levels, even if the true transaction was a single sale. The unweighted model would double-count these observations while the weighted model is "invariant" to that allocation.  Whether a 100 pounds transaction enters as 1 row, 10 rows, or 100 rows, has no impact on the weighted model but would affect the unweighted model results. 

The estimating equation for the hedonic model is 
\begin{align}
p_{id}&= \sum_{k=1}^K a_{k} Market Category_{id} + \sum_{g=1}^G b_{k} Gear_{id} + \label{HedonicEstimatingEquation}\\
 &+ d Daily Landings_{d} +   e Daily Landings^2_{d}  + f Live_{id} + \varepsilon_i\notag
\end{align}

We also include a set of factor variables for state, year, and month; these are not interacted for simplicity. We estimate equation \ref{HedonicEstimatingEquation} with a full set of factor variables for Market Category and exclude 1 level from all other factors to avoid perfect linearity.  Equation \ref{HedonicEstimatingEquation} is estimated by Ordinary Least Squares. While other more sophisticated models could be used here, OLS suffices for these exploratory purposes.   The omitted levels are Trawl Gear, Rhode Island, 2013, and January.  The coefficients associated with the market categories can be interpreted as the conditional average price per pound of landings in that category that were caught by trawl gear and landed in Rhode Island in January of 2013.  Coefficients for the other factor variables can be interpreted as premia or discounts relative to the omitted factor.

Table \ref{HedonicTableA} contains some of the more interesting results. There was 46.9 million pounds (21,280mt) of Black Sea Bass landed and sold in nearly 572,000 transactions during the 2013-2025 time period.  We prefer the weighted specification, which has a more intuitive interpretation. Prices of Jumbo Black Sea Bass caught by trawl gear in Rhode Island in January of 2013 average \$13.08 per kg, while Small Black Sea Bass in the same area are priced at \$7.46 per kg.  Unclassified Black Sea Bass were priced between the Medium and Large market categories, suggesting that average size of  Unclassified fish lies between those two categories. All of the gear categories receive premia relative to trawl caught fish.  Live fish receives a slight premia relative to Round fish. Increases in Total landings cause slight declines in prices; this occurs at a decreasing rate. 

Figure \@ref(fig:HedonicMarginals) shows the results for the month, year, and state factor variables. The Year dummies indicate a substantial decrease in the real price of Black Sea Bass over the 2013-2025 period.  Prices also vary by year, month, and state. 

This exploratory analysis serves three purposes. First, it confirms that prices vary by market category; this effect persists even after conditioning out effects of gear, landing state, year, month, and aggregate supply effects.  Second, simple hedonic model provides further evidence of a monotone relationship between size and price. Third, it suggests explanatory factors that belong in the Random Forest model. Without including these covariates, we risk confounding the effects of these explanatory variables on price.   


```{r hedonicsummary,  eval=TRUE, echo=FALSE}
# this comes out of data_prep_ml.Rmd. It's 
bsb_original<-readr::read_rds(file=here("data_folder","main","commercial", glue("BSB_original_combined_dataset{data_vintage_string}.Rds")))
lbs_per_kg<-2.20462

bsb_original<-bsb_original %>%
  filter(mark_in==1) %>%
  mutate(lndkg=lndlb/lbs_per_kg)%>%
  mutate(pricekgR_CPI=valueR_CPI/lndkg)

bsb_original<-bsb_original %>%
  group_by(dlr_date) %>%
  mutate(daily_landings_mt=sum(lndkg, na.rm=TRUE)/1000)




my_selection   <- c("pricekgR_CPI", "daily_landings_mt")
more_variables <- c("market_desc", "mygear", "grade_desc", "state")

# Readable labels for continuous variables
continuous_labels <- c(
  pricekgR_CPI     = "Price per kilogram (Real)",
  daily_landings_mt = "Landings (mt)"
)
# Readable labels for continuous variables
factor_labels <- c(
  market_desc     = "Market Category",
  mygear          = "Gear",
  grade_desc      = "Round/Live",
  state           = "State"
  )



# -------------------------------------------------------------------
# Helper: weighted mean and SD
# -------------------------------------------------------------------
wtd_mean <- function(x, w) {
  keep <- !is.na(x) & !is.na(w)
  sum(x[keep] * w[keep]) / sum(w[keep])
}

wtd_sd <- function(x, w) {
  keep <- !is.na(x) & !is.na(w)
  xw   <- wtd_mean(x[keep], w[keep])
  sqrt(sum(w[keep] * (x[keep] - xw)^2) / sum(w[keep]))
}

# -------------------------------------------------------------------
# 1. weights and obs 
# -------------------------------------------------------------------
n_row <- tibble(
  variable = "N",
  level    = "",
  Weighted   = formatC(sum(bsb_original$weighting),  format = "f", digits = 0, big.mark = ","),
  Unweighted = formatC(nrow(bsb_original),  format = "f", digits = 0, big.mark = ",")
)

# -------------------------------------------------------------------
# 2. Continuous variables: mean + SD row pairs
# -------------------------------------------------------------------
cont_rows <- purrr::map_dfr(my_selection, function(v) {
  x <- bsb_original[[v]]
  w <- bsb_original[["weighting"]]

  mean_w  <- wtd_mean(x, w)
  sd_w    <- wtd_sd(x, w)
  mean_uw <- mean(x, na.rm = TRUE)
  sd_uw   <- sd(x,   na.rm = TRUE)

  bind_rows(
    tibble(
      variable   = continuous_labels[v],
      level      = "",
      Weighted   = formatC(mean_w,  format = "f", digits = 2, big.mark = ","),
      Unweighted = formatC(mean_uw, format = "f", digits = 2, big.mark = ",")
    ),
    tibble(
      variable   = "",
      level      = "",
      Weighted   = paste0("(", formatC(sd_w,  format = "f", digits = 2,
                                       big.mark = ","), ")"),
      Unweighted = paste0("(", formatC(sd_uw, format = "f", digits = 2,
                                       big.mark = ","), ")")
    )
  )
})

# -------------------------------------------------------------------
# 3. Factor variables: proportion per level
# -------------------------------------------------------------------
factor_rows <- purrr::map_dfr(more_variables, function(v) {
  x <- bsb_original[[v]]
  w <- bsb_original[["weighting"]]
  lvls <- sort(unique(x[!is.na(x)]))

  header <- tibble(
    variable   = factor_labels[v],
    level      = "",
    Weighted   = "",
    Unweighted = ""
  )

  level_rows <- purrr::map_dfr(lvls, function(l) {
    idx  <- x == l & !is.na(x)
    prop_w  <- sum(w[idx], na.rm = TRUE) / sum(w[!is.na(x)], na.rm = TRUE)
    prop_uw <- mean(idx, na.rm = TRUE)

    tibble(
      variable   = "",
      level      = as.character(l),
      Weighted   = formatC(prop_w,  format = "f", digits = 3),
      Unweighted = formatC(prop_uw, format = "f", digits = 3)
    )
  })

  bind_rows(header, level_rows)
})

# -------------------------------------------------------------------
# 4. Bind all sections together
# -------------------------------------------------------------------
summary_table <- bind_rows(cont_rows, factor_rows, n_row)

kable(
  summary_table,
  format    = "latex",
  booktabs  = TRUE,
  caption   = "Weighted and unweighted summary statistics. Continuous variables
              display means with standard deviation in parentheses. Factor 
              variables display proportion of observations in each level. 
              Weights are pounds landed per Transaction'",
  col.names = c("Variable", "Level", "Weighted", "Unweighted"),
  escape    = TRUE,
  label     = "bsb-hedonic-summary",
  align     = c("l", "l", "r", "r")
) %>%
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 9
  ) %>%
  column_spec(1, width = "4cm", bold = TRUE) %>%
  column_spec(2, width = "3cm") %>%
  column_spec(3:4, width = "2.5cm")
```

\clearpage


<!---This will read a .tex file.  

The hedonic model is estimated by 
global in_string 2026-03-03
do "$analysis_code/00_analysis_wrapper.do"

 --->
\input{../results/hedonic_tableA.tex}

\clearpage

```{r HedonicMarginals,echo=FALSE,out.width="48%", fig.show='hold', fig.align='center',fig.cap="The marginal effects of month, year, and state in the hedonic model capture the prevailing price premium or discount for these categorical factors relative to the omittied baseline."}
knitr::include_graphics(here("results",
                              c("HedonicMonth.png",
                              "HedonicYear.png",
                              "HedonicState.png")
                           )
                        )
```

\clearpage