---
title: "Black Sea Bass: Calibrate and Validate"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: noclusterTEST
    choices:
    - nocluster
    - noclusterTEST
---

# Summary and Housekeeping


Calibration and Validation

1. Read in the finalized model 
2. Calibrate probabilities using a 'testing sample'
3. Validate on a validation sample. 

<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	Read in outputs of machine learning models 
* Inputs: These three can be adjusted by changing the "modeltype"
*   - Data file
*   - tuning results file
*   - final results file


* Outputs:
*   - Graphs of tuning quality and final model fit
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


**********************************************************************

 
 --->
```{r global_options, include=FALSE}

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())
conflicts_prefer(vip::vi)

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/calibration_and_validation.Rmd")

#handle model patterns
source(here("R_code","analysis","helpers","modeltype_patterns.R"))




#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)

# My "final fit" bombed out because I didn't have drive space.  
# finalfit_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{final_pattern}*Rds")))
# finalfit_vintage<-gsub(final_pattern,"",finalfit_vintage)
# finalfit_vintage<-gsub(".Rds","",finalfit_vintage)
# finalfit_vintage<-max(finalfit_vintage)

finalfit_vintage<-tuning_vintage

vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)



```


# Load in the final model

read in the final model, extract the recipe and the predictors, extract the predictions

```{r load_final_model}
# # this was created with data_prep_ml.Rmd
data_split<-readr::read_rds(file=here("results","ranger",glue("{data_pattern}{data_vintage_string}.Rds")))
# Onlyneed the test and validation 
# train_data <- training(data_split)
validation_data <- validation(data_split)
test_data <- testing(data_split)
rm(data_split)


final_fit<-read_rds(file=here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))


# What is the workflow
final_workflow<-extract_workflow(final_fit)

# And the recipe
extract_recipe(final_fit)

preds<-final_fit$.workflow[[1]]$pre$actions$recipe$recipe$var_info %>%
 dplyr::filter(role=="predictor")

num_preds<-nrow(preds)

final_predictions<-final_fit %>%
  collect_predictions()

# pull the probability names
prob_names<-colnames(final_predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


# look at the fit metrics
final_fit_metrics<-final_fit %>%
  collect_metrics()
final_fit_metrics

```

Log loss of the final fit is slightly better than that of the average of the individual folds. This is normal/reasonable, since it uses more data than each of the individual folds.  It's not so different that alarm bells go off.


## Validation Diagnostics

Look at the VIP, ROC, and PR curves

```{r VIP}
# Look at variable importance. I want to see all the predictors.
# The scale is different from previous versions, because I switched the variable importance method in the Workflow/recipe to "Permutation"
vi_data <- final_fit %>%
  extract_fit_parsnip() %>%
  vi(method = "model") %>%          # pulls ranger's permutation importance
  #slice_max(Importance, n = 20) %>% # top 20 only at single-column width
  mutate(
    Variable = forcats::fct_reorder(Variable, Importance),
    Variable = forcats::fct_recode(Variable,
      "Price Difference Jumbo" = "Price_Diff_J",
      "Price Difference Large" = "Price_Diff_L",
      "Price Difference Medium" = "Price_Diff_M",
      "Dealer Propensity Small" = "LagSharePoundsSmall",
      "Dealer Propensity Medium" = "LagSharePoundsMedium",
      "Dealer Propensity Large" = "LagSharePoundsLarge",
      "Dealer Propensity Jumbo" = "LagSharePoundsJumbo",
      "Year" = "year",
      "Stockarea Catch Jumbo" = "MA7_StockareaQJumbo",
      "Stockarea Catch Large" = "MA7_StockareaQLarge",
      "Stockarea Catch Medium" = "MA7_StockareaQMedium",
      "Stockarea Catch Small" = "MA7_StockareaQSmall",
      "State Catch Jumbo" = "MA7_StateQJumbo",
      "State Catch Large" = "MA7_StateQLarge",
      "State Catch Medium" = "MA7_StateQMedium",
      "State Catch Small" = "MA7_StateQSmall",
      "Gear Catch Jumbo" = "MA7_gearQJumbo",
      "Gear Catch Large" = "MA7_gearQLarge",
      "Gear Catch Medium" = "MA7_gearQMedium",
      "Gear Catch Small" = "MA7_gearQSmall",
      "Transaction Weight" = "lndlb",
      "Stockarea Trips" = "MA7_stockarea_trips",
      "State Trips" = "MA7_state_trips"
      )
    )



p_vip <- ggplot(vi_data %>% slice_max(Importance, n=20), aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1B6CA8", width = 0.7) +
  geom_vline(xintercept = 0, colour = "grey20", linewidth = 0.3) +
  scale_x_continuous(
    name   = "Mean Decrease in Accuracy",
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_y_discrete(name = NULL) +
  theme_bw(base_size = 9) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    axis.text.y        = element_text(size = 7, colour = "grey20"),
    axis.text.x        = element_text(size = 7, colour = "grey20"),
    axis.title.x       = element_text(size = 8),
    plot.margin        = margin(4, 6, 4, 4, "pt")
  )
p_vip
# --- 3. Save at ICES JMS single-column specification ---
ggsave(
  here("results", "ranger", "final",
       glue("vip{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_vip,
  width  = 84,
  height = 110,    # 20 bars fit cleanly; adjust in 5mm increments if needed
  units  = "mm",
  device = cairo_pdf
)



p_vip <- ggplot(vi_data,  aes(x = Importance, y = Variable)) +
  geom_col(fill = "#1B6CA8", width = 0.7) +
  geom_vline(xintercept = 0, colour = "grey20", linewidth = 0.3) +
  scale_x_continuous(
    name   = "Mean Decrease in Accuracy",
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_y_discrete(name = NULL) +
  theme_bw(base_size = 9) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    axis.text.y        = element_text(size = 7, colour = "grey20"),
    axis.text.x        = element_text(size = 7, colour = "grey20"),
    axis.title.x       = element_text(size = 8),
    plot.margin        = margin(4, 6, 4, 4, "pt")
  )

# --- 3. Save at ICES JMS single-column specification ---
ggsave(
  here("results", "ranger", "final",
       glue("vipFULL{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_vip,
  width  = 84,
  height = 110,    # 20 bars fit cleanly; adjust in 5mm increments if needed
  units  = "mm",
  device = cairo_pdf
)



```
Variable importance measures are the Prices and the Dealer proporption variables. 


```{r PR_Curve}
# pull the probability names
prob_names<-colnames(final_predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)



final_pr<-final_predictions %>%
  pr_curve(market_desc,  any_of(prob_names)) %>%
  autoplot()
ggsave(here("results","ranger","final",glue("pr_curve{modeltype}{finalfit_vintage}.png")), plot=final_pr)
```

# Validation and Calibration 

Use the Validation set to (1 assess how well the model fits) and 2. Calibrate if I have systematic under or over prediction.


```{r Validation_predictions}
# Apply the Final fit to the validation dataset. 

validation_data <- augment(final_workflow,validation_data)

#The validation plots indicate characteristic issues with probability trees. All models are underconfident. Groups of observations with a 75% predicted probability of Jumbo have a true probability of Jumbo of almost 90%.  Groups of observations with a predicted probability of ~15% probabilty of Jumbo have a true probability closer to 10%.

# Calibration should fix this.

cal_gg <- validation_data %>%
  cal_plot_windowed(
    truth          = market_desc,
    estimate       = c(.pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small),
    step_size      = 0.025,
    include_points = FALSE
  )

# Extract data from the ggplot internals
cal_data <- cal_gg$data



# --- 2. Build publication-ready faceted calibration plot ---
p_cal <- ggplot(cal_data,
                aes(x = predicted_midpoint, y = event_rate)) +
  # perfect calibration reference line
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  # confidence band
  geom_ribbon(aes(ymin = lower, ymax = upper),
              fill = "#1B6CA8", alpha = 0.15) +
  # calibration curve
  geom_line(colour = "#1B6CA8", linewidth = 0.8) +
  facet_wrap(~ market_desc, ncol = )+
  scale_x_continuous(
    name   = "Mean Predicted Probability",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Observed Event Rate",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
p_cal
# --- 3. Save at ICES JMS double-column specification ---
ggsave(
  here("results", "ranger", "final",
       glue("cal_windowed_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_cal,
  width  = 84,
  height = 88,     # suits 4 panels in one row; increase to 130 for 2x2 layout
  units  = "mm",
  device = cairo_pdf
)


```


Experiment with different methods of calibration. Beta fits a 3 parameter beta model (2 shape, 1 location), performing 4 distinct 1 vs all models and then re-normalalizing the results.  Iso fits an isotonic regression (4 separate 1 vs all again).  cal_estimate_multinom fits logistic splines and seemed to work better than the version that just fit a multinomial logistic.

```{r calibration_models}

calibrate_beta <- cal_estimate_beta(validation_data, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)

calibrate_iso <- cal_estimate_isotonic(validation_data, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)

calibrate_multinom <- cal_estimate_multinomial(validation_data, 
                                        truth=market_desc, 
                                        estimate=.pred_Jumbo:.pred_Small,
                                        smooth=TRUE)

# weights <- validation_data$lndlb
# 
# calibrate_multinomW <- cal_estimate_multinomial(validation_data, 
#                                         truth=market_desc, 
#                                         estimate=.pred_Jumbo:.pred_Small,
#                                         smooth=FALSE, parameters=NULL, weights=lndlb)


```





```{r apply_calibration}
# This is fine for testing out things, but the final judging is done on the 'testing' dataset

validation_data_isocalib_applied <-
  validation_data %>%
  cal_apply(calibrate_iso)

validation_data_betacalib_applied <-
  validation_data %>%
  cal_apply(calibrate_beta)

validation_data_multicalib_applied <-
  validation_data %>%
  cal_apply(calibrate_multinom)


# subesequent testing shows this is the best calibration, so we save it.
write_rds(calibrate_iso, file=here("results","ranger","final",glue("calibrate_iso_{modeltype}{finalfit_vintage}.Rds")))


```


```{r plot_iso_calib}
# Pub read isotonic calibration

# Calibration should fix this.
cal_gg <- validation_data_isocalib_applied %>%
  cal_plot_windowed(
    truth          = market_desc,
    estimate       = c(.pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small),
    step_size      = 0.025,
    include_points = FALSE
  )

# Extract data from the ggplot internals
cal_data <- cal_gg$data



# --- 2. Build publication-ready faceted calibration plot ---
iso_applied_window <- ggplot(cal_data,
                aes(x = predicted_midpoint, y = event_rate)) +
  # perfect calibration reference line
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  # confidence band
  geom_ribbon(aes(ymin = lower, ymax = upper),
              fill = "#1B6CA8", alpha = 0.15) +
  # calibration curve
  geom_line(colour = "#1B6CA8", linewidth = 0.8) +
  facet_wrap(~ market_desc, ncol = )+
  scale_x_continuous(
    name   = "Mean Predicted Probability",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Observed Event Rate",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
iso_applied_window
# --- 3. Save at ICES JMS double-column specification ---
ggsave(
  here("results", "ranger", "final",
       glue("cal_isovalidate_window{modeltype}{finalfit_vintage}.pdf")),
  plot   = iso_applied_window,
  width  = 84,
  height = 88,     # suits 4 panels in one row; increase to 130 for 2x2 layout
  units  = "mm",
  device = cairo_pdf
)
```

```{r other_calibs}
#Multi looks a little rough
multi_applied_window<-cal_plot_windowed(validation_data_multicalib_applied,
                  truth = market_desc, step_size = 0.05, include_points =FALSE)
multi_applied_window
ggsave(here("results","ranger","final",glue("cal_multicalibrate_window{modeltype}{finalfit_vintage}.png")), plot=multi_applied_window)



beta_applied_window<-cal_plot_windowed(validation_data_betacalib_applied,
                  truth = market_desc, step_size = 0.05, include_points =FALSE)
ggsave(here("results","ranger","final",glue("cal_betavalidate_window{modeltype}{finalfit_vintage}.png")), plot=beta_applied_window)



# Compute the Calibration Loss (from Ferrer and Ramos "Evaluating Posterior Probabilities")
# (ESPR_raw - ESPR_cal)/ESPR_raw

ESPR_raw<-validation_data %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_caliso<-validation_data_isocalib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_calbeta<-validation_data_betacalib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_calmulti<-validation_data_multicalib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

# the difference between these metrics is the calibration loss (A little weird). But basically, the "goodness of fit metric" improves when I apply a calibration. the multinomial model doesn't work so well. But Iso and Beta do. Iso is slightly better, so we will use that.
ESPR_raw
ESPR_caliso
ESPR_calbeta
ESPR_calmulti

CalLoss<-ESPR_raw-ESPR_caliso
rCalLoss<-100*CalLoss/ESPR_raw

CalLoss
rCalLoss
```







```{r ROC_Curve, eval=FALSE}

# Reciever operator curve (True positive vs False Positive) for the final model evaluated on the validation data
final_roc<-validation_data_isocalib_applied %>%
  roc_curve(truth=market_desc, 
            any_of(prob_names),
            case_weights=weighting)  %>%
  autoplot()
ggsave(here("results","ranger","final",glue("validation_roc_curve{modeltype}{finalfit_vintage}.png")), plot=final_roc)



final_roc_nocal<-validation_data %>%
  roc_curve(truth=market_desc, 
            any_of(prob_names),
            case_weights=weighting)  %>%
  autoplot()
ggsave(here("results","ranger","final",glue("validation_uncal_roc_curve{modeltype}{finalfit_vintage}.png")), plot=final_roc_nocal)


```





```{r pub_ready_ROC_uncalibrated}
# --- 1. Compute per-class ROC curve data ---
roc_data <- validation_data %>%
  roc_curve(
    truth = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small,
    case_weights=weighting
  )


roc_dataUW <- validation_data %>%
  roc_curve(
    truth = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small
  )



# --- 2. Compute per-class AUC for annotation ---
auc_dataW <- validation_data %>%
  roc_auc(
    truth     = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small,
    case_weights=weighting,
    estimator = "macro"
  )


auc_dataUW <- validation_data %>%
  roc_auc(
    truth     = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small,
    estimator = "macro"
  )

# --- 3. Colour palette — four classes ---
class_colours <- c(
  "Jumbo"  = "#1B6CA8",   # deep blue
  "Large"  = "#E05C2A",   # burnt orange
  "Medium" = "#2E8B57",   # sea green
  "Small"  = "#7B3F9E"    # purple
)

# --- 4. Build the figures ---
p_facet <- ggplot(roc_data,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level), linewidth = 0.8, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
p_facet
# --- 5. Save at ICES JMS specifications ---
ggsave(here("results","ranger","final",glue("validation_nocal_roc_curve{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facet,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)




# --- 6. Build the unweighted figure ---
p_facetUW <- ggplot(roc_dataUW,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level), linewidth = 0.8, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
p_facetUW
ggsave(here("results","ranger","final",glue("validation_nocal_UWroc_curve{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facetUW,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)



```







```{r pub_ready_ROC_calibrated}
# --- 1. Compute per-class ROC curve data ---
roc_data <- validation_data_isocalib_applied %>%
  roc_curve(
    truth = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small,
    case_weights=weighting
  )


roc_dataUW <- validation_data_isocalib_applied %>%
  roc_curve(
    truth = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small
  )



# --- 2. Compute per-class AUC for annotation ---
auc_data <- validation_data_isocalib_applied %>%
  roc_auc(
    truth     = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small,
    case_weights=weighting,
    estimator = "macro"
  )

# --- 3. Colour palette — four classes ---
class_colours <- c(
  "Jumbo"  = "#1B6CA8",   # deep blue
  "Large"  = "#E05C2A",   # burnt orange
  "Medium" = "#2E8B57",   # sea green
  "Small"  = "#7B3F9E"    # purple
)

# --- 4. Build the figures ---
p_facet <- ggplot(roc_data,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level), linewidth = 0.8, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )

p_facet
# --- 5. Save at ICES JMS specifications ---
ggsave(here("results","ranger","final",glue("validation_calibrated_roc_curve{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facet,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)




# --- 6. Build the unweighted figure ---
p_facetUW <- ggplot(roc_dataUW,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level), linewidth = 0.8, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
p_facetUW
ggsave(here("results","ranger","final",glue("validation_calibrated_UWroc_curve{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facetUW,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)



```



















```{r final_model_calibration}

# Predict Out of Sample on the testing holdout.


test_data <- augment(final_workflow,test_data)

############ How well calibrated is the model? 
# A: Better when calibrated - same pattern as training data.
# Plot the "raw" predictions
uncalibrated<-cal_plot_windowed(test_data, truth = market_desc, step_size = 0.05,include_points=FALSE)

#ggsave(here("results","ranger","final",glue("validation_raw_{modeltype}{finalfit_vintage}.png")), plot=uncalibrated)
#write_rds(uncalibrated, file=here("results","ranger","final",glue("validation_raw_{modeltype}{finalfit_vintage}.Rds")))


test_data_calibration_applied <-
  test_data %>%
  cal_apply(calibrate_iso)

# Plot the calibrated predictions
# need to pub-ready this.







calibrated<-cal_plot_windowed(test_data_calibration_applied, truth = market_desc, step_size = 0.05,include_points=FALSE)

calibrated

# Pub Ready Calibration
# Extract data from the ggplot internals
cal_data <- calibrated$data



# --- 2. Build publication-ready faceted calibration plot ---
p_cal <- ggplot(cal_data,
                aes(x = predicted_midpoint, y = event_rate)) +
  # perfect calibration reference line
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  # confidence band
  geom_ribbon(aes(ymin = lower, ymax = upper),
              fill = "#1B6CA8", alpha = 0.15) +
  # calibration curve
  geom_line(colour = "#1B6CA8", linewidth = 0.8) +
  facet_wrap(~ market_desc, ncol = )+
  scale_x_continuous(
    name   = "Mean Predicted Probability",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Observed Event Rate",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
p_cal
# --- 3. Save at ICES JMS double-column specification ---
ggsave(
  here("results", "ranger", "final",
       glue("cal_windowed_testing{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_cal,
  width  = 84,
  height = 88,     # suits 4 panels in one row; increase to 130 for 2x2 layout
  units  = "mm",
  device = cairo_pdf
)





# Pub Ready Calibration
# Extract data from the ggplot internals
uncal_data <- uncalibrated$data



# --- 2. Build publication-ready faceted calibration plot ---
p_uncal <- ggplot(uncal_data,
                aes(x = predicted_midpoint, y = event_rate)) +
  # perfect calibration reference line
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  # confidence band
  geom_ribbon(aes(ymin = lower, ymax = upper),
              fill = "#1B6CA8", alpha = 0.15) +
  # calibration curve
  geom_line(colour = "#1B6CA8", linewidth = 0.8) +
  facet_wrap(~ market_desc, ncol = )+
  scale_x_continuous(
    name   = "Mean Predicted Probability",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Observed Event Rate",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
p_uncal
# --- 3. Save at ICES JMS double-column specification ---
ggsave(
  here("results", "ranger", "final",
       glue("uncal_windowed_testing{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_uncal,
  width  = 84,
  height = 88,     # suits 4 panels in one row; increase to 130 for 2x2 layout
  units  = "mm",
  device = cairo_pdf
)



```




```{r pub_ready_test_ROC_calibrated}
# --- 1. Compute per-class ROC curve data ---
roc_data <- test_data_calibration_applied %>%
  roc_curve(
    truth = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small,
    case_weights=weighting
  )


roc_dataUW <- test_data_calibration_applied %>%
  roc_curve(
    truth = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small
  )



# --- 2. Compute per-class AUC for annotation ---
auc_data <- test_data_calibration_applied %>%
  roc_auc(
    truth     = market_desc,
    .pred_Jumbo, .pred_Large, .pred_Medium, .pred_Small,
    case_weights=weighting,
    estimator = "macro"
  )

# --- 3. Colour palette — four classes ---
class_colours <- c(
  "Jumbo"  = "#1B6CA8",   # deep blue
  "Large"  = "#E05C2A",   # burnt orange
  "Medium" = "#2E8B57",   # sea green
  "Small"  = "#7B3F9E"    # purple
)

# --- 4. Build the figures ---
p_facet <- ggplot(roc_data,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level), linewidth = 0.8, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )

p_facet
# --- 5. Save at ICES JMS specifications ---
ggsave(here("results","ranger","final",glue("testing_calibrated_roc_curve{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facet,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)




# --- 6. Build the unweighted figure ---
p_facetUW <- ggplot(roc_dataUW,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level), linewidth = 0.8, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
p_facetUW
ggsave(here("results","ranger","final",glue("testing_calibrated_UW_ROC_curve{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facetUW,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)



```










```{r calibration_gains}

#cal_plot_breaks(test_data_calibration_applied, truth = market_desc, num_breaks=20,include_points=FALSE)

# Compute the Calibration Loss (from Ferrer and Ramos "Evaluating Posterior Probabilities")
# (ESPR_raw - ESPR_cal)/ESPR_raw

ESPR_raw<-test_data %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)


ESPR_cal<-test_data_calibration_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_raw
ESPR_cal

CalLoss<-ESPR_raw-ESPR_cal
rCalLoss<-100*CalLoss/ESPR_raw

#  The iso calibration improves model fit a bit. 

CalLoss
rCalLoss
```

# Calibrated Predictions

In this section, I predict out of sample on my 'test' dataset with the calibration applied 


```{r extract_predictions}
### pull a few categories, compute a prob*lndlb variable, sum 


# keep just a few columns
test_data_calibration_applied<-test_data_calibration_applied %>%
  select(c( weighting, market_desc, lndlb, stockarea, year, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the landed pounds to get predicted pounds in each class. 
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    test_data_calibration_applied[[paste0("pred_", l)]] <- test_data_calibration_applied[[paste0(".pred_", l)]] * test_data_calibration_applied$lndlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}
# Sum up the predictions by year, market category, and stockarea. 
test_predictions<-test_data_calibration_applied %>%
   group_by(stockarea, year,market_desc) %>%
  select(-.pred_class) %>%
  mutate(transactions=1) %>%
  summarise(across(c(starts_with(c("pred_",".pred_")), "lndlb", "transactions"), \(x) sum(x, na.rm = TRUE))) %>%
   ungroup()
test_predictions$modeltype<-modeltype

test_predictions<-test_predictions %>%
  mutate(market_desc=fct_relevel(market_desc,"Jumbo","Large","Medium","Small")) 

write_rds(test_predictions, file=here("results","ranger",glue("aggregate_test_predictions_{modeltype}{finalfit_vintage}.Rds")))

#test_predictions<-test_predictions %>%
#  dplyr::filter(year %in% c(2021,2022,2023,2024))

```

## Probabalistic Predictions of the class of each transaction

Here are the predictions (transactions and mt) for the hold-out sample, using the probability calibrations.

```{r topline_validation}

aggregate_transactions<-test_predictions %>%
    summarise(across(c(starts_with(".pred_")), sum)) %>%
  pivot_longer(cols=starts_with(".pred_"), names_to="market_desc", names_prefix=".pred_",values_to="ObsPredicted")


aggregate_test_predictions<-test_predictions %>%
    summarise(across(c(starts_with("pred_")), sum)) %>%
  pivot_longer(cols=starts_with("pred_"), names_to="market_desc", names_prefix="pred_",values_to="predicted")


true<-test_predictions %>%
    group_by(market_desc) %>%
    summarise(lndlb=sum(lndlb),
              transactions=sum(transactions))

Testmkt_preds<-aggregate_transactions %>%
  left_join(true, by=join_by(market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  left_join(aggregate_test_predictions, by=join_by(market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted,
         predicted_transactions=ObsPredicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01),
         `Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )


transaction_predictions<-Testmkt_preds %>%
  select(market_desc,transactions, predicted_transactions, `Transaction error (%)`)


knitr::kable(transaction_predictions, caption='Calibrated Predictions on the 15% Validation Sample (transaction count).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```

```{r weighted_predictions}
weighted_predictions<-Testmkt_preds %>%
  select(market_desc,true_mt, predicted_mt, `mt error (%)`)

knitr::kable(weighted_predictions, caption='Calibrated Predictions on the 15% Validation Sample (mt).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```


Here are the predictions (transactions and mt) for the hold-out sample, without the probability calibrations.

```{r aggregate_test2_predictions}
# Subset data, construct predicted pounds, and aggregate


### pull a few categories, compute a prob*lndlb variable, sum 


# keep just a few columns
test_data_nocalibration_applied<-test_data%>%
  select(c( weighting, market_desc, lndlb, stockarea, year, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the landed pounds to get predicted pounds in each class. 
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    test_data_nocalibration_applied[[paste0("pred_", l)]] <- test_data_nocalibration_applied[[paste0(".pred_", l)]] * test_data_nocalibration_applied$lndlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}
# Sum up the predictions by year, market category, and stockarea. 
nocal_predictions<-test_data_nocalibration_applied%>%
   group_by(stockarea, year,market_desc) %>%
  select(-.pred_class) %>%
  mutate(transactions=1) %>%
  summarise(across(c(starts_with(c("pred_",".pred_")), "lndlb", "transactions"), \(x) sum(x, na.rm = TRUE))) %>%
   ungroup()
nocal_predictions$modeltype<-modeltype

nocal_predictions<-nocal_predictions%>%
  mutate(market_desc=fct_relevel(market_desc,"Jumbo","Large","Medium","Small")) 

write_rds(nocal_predictions, file=here("results","ranger",glue("aggregate_nocal_predictions_{modeltype}{finalfit_vintage}.Rds")))

```







```{r no_cal_validations}

aggregate_nocal_transactions<-nocal_predictions%>%
    summarise(across(c(starts_with(".pred_")), sum)) %>%
  pivot_longer(cols=starts_with(".pred_"), names_to="market_desc", names_prefix=".pred_",values_to="ObsPredicted")


aggregate_nocal_weight_predictions<-nocal_predictions %>%
    summarise(across(c(starts_with("pred_")), sum)) %>%
  pivot_longer(cols=starts_with("pred_"), names_to="market_desc", names_prefix="pred_",values_to="predicted")


true<-test_data%>%
    mutate(transactions=1) %>%
    group_by(market_desc) %>%
    summarise(lndlb=sum(lndlb),
              transactions=sum(transactions))

nocalibrate_results<-aggregate_nocal_weight_predictions%>%
  left_join(true, by=join_by(market_desc==market_desc))

nocalibrate_results<-nocalibrate_results%>%
  left_join(aggregate_nocal_transactions, by=join_by(market_desc==market_desc))

nocalibrate_results<-nocalibrate_results%>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted,
         predicted_transactions=ObsPredicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01),
         `Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )
  

transactions_nocalibrate_predictions<-nocalibrate_results %>%
  select(market_desc,transactions, predicted_transactions, `Transaction error (%)`)


knitr::kable(transactions_nocalibrate_predictions, caption='Uncalibrated Predictions on the 15% Validation Sample (transactions).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```


```{r weighted_no_calibrate_predictions}
weights_nocalibrate_predictions<-nocalibrate_results %>%
  select(market_desc,true_mt, predicted_mt, `mt error (%)`)
knitr::kable(weights_nocalibrate_predictions, caption='Uncalibrated Predictions on the 15% Validation Sample (mt).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```









## Lets take a look at the validation predictions by stockarea



```{r Validation_Weight_Predictions_by_stockarea}

# Aggregate to the stockarea- market category

aggregate_test_predictions2<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(across(c(starts_with(".pred_"),lndlb), sum)) %>%
  arrange(stockarea,market_desc)

# true landings by stockarea and market_category
trueTest<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(lndlb=sum(lndlb))


# aggregate predictions, in pounds, by stockarea and market_category
Tp1<-test_predictions %>%
  group_by(stockarea) %>%
    summarise(across(c(starts_with("pred_")), sum))

Tp2<-Tp1  %>%
  pivot_longer(cols=!stockarea, names_to="market_desc", names_prefix="pred_",values_to="predicted")


Testmkt_preds<-Tp2 %>%
  left_join(trueTest, by=join_by(stockarea==stockarea, market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01) )


knitr::kable(Testmkt_preds, caption='Calibrated Predictions on the 15% Validation Sample (mt). Disaggregated to Stock area',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```




```{r Validation_Transaction_Predictions_by_stockarea}


# true landings by stockarea and market_category
trueTest<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(transactions=sum(transactions))


# aggregate predictions, in transactions, by stockarea and market_category
Tp1<-test_predictions %>%
  group_by(stockarea) %>%
    summarise(across(c(starts_with(".pred_")), sum))

Tp2<-Tp1  %>%
  pivot_longer(cols=!stockarea, names_to="market_desc", names_prefix=".pred_",values_to="predicted")


Testmkt_preds<-Tp2 %>%
  left_join(trueTest, by=join_by(stockarea==stockarea, market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  rename(predicted_transactions=predicted) %>%
  mutate(`Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )


knitr::kable(Testmkt_preds, caption='Calibrated Predictions on the 15% Validation Sample (observations). Disaggregated to Stock area',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  
```



# Things that didn't work

I tried to unweight and the fit another calibration. It didn't work too well.


```{r explore_validations, eval=FALSE}
# This section 'unweights' the validation data and then tries to fit a calibration.
# It doesn't work too well, in part because the initial cal_plot is non-monotone. 
# The observation-level model is textbook smooth, the unweighted one is not. 
# This makes the parametric (beta, multi SMOOTH=FALSE) fit poorly
# The multi Smooth and isotonic "should" fit better. 
# The isotonic takes forever to run, because it has to do alot of comparisons.

validation_expanded <- validation_data %>%
  uncount(weights = weighting)

# this dataset is massive. take a 5% sample, just to see what is going on.

validation_sub<-initial_split(validation_expanded,prop=0.90)
tt<-testing(validation_sub)


tt %>%
    cal_plot_windowed(truth = market_desc,step_size = 0.05)

tt %>%
  cal_plot_breaks(truth = market_desc, step_size = 0.025)


calibrate_betaTT <- cal_estimate_beta(tt, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)

calibrate_multiTT <- cal_estimate_multinomial(tt, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=TRUE)

calibrate_isoTT <- cal_estimate_isotonic(tt, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)


WT_data_calib_applied<-  tt %>%
  cal_apply(calibrate_multiTT)


WT_data_calib_applied %>%
    cal_plot_windowed(truth = market_desc,step_size = 0.05)



W_ESPR_raw<-tt %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

W_ESPR_multi<-WT_data_calib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)


CalLoss<-W_ESPR_raw-W_ESPR_multi
rCalLoss<-100*CalLoss/W_ESPR_raw

CalLoss
rCalLoss


# calibrate_iso <- cal_estimate_isotonic(validation_expanded, 
#                                    truth=market_desc, 
#                                    estimate=.pred_Jumbo:.pred_Small,
#                                    smooth=FALSE)





validation_exp_data_isocalib_applied <-
  validation_expanded %>%
  cal_apply(calibrate_iso)

validation_exp_data_isocalib_applied %>%
  cal_plot_breaks(truth = market_desc, step_size = 0.025)


```




<!---
\newpage
--->
# References
<div id="refs"></div>





# Appendix{-}