---
title: 'Understanding the Impacts of state-level landings limits'
author: 
  - Min-Yang Lee, NOAA Fisheries
  - Anna Birkenbach, University of Delaware
date: "`r format(Sys.time(), '%B %d, %Y')`" 
documentclass: article
csl: "../ajae_mod.csl"
abstract: ""
output:
  word_document: null
  html_document: null
  pdf_document: 
    includes:
      in_header: "../preamble-latex.tex"
    keep_tex: yes
    pandoc_args: --pdf-engine=pdflatex
    number_sections: true
fontsize: 12pt
bibliography: "../../library.bib"
---


 <!---- 
Summary and Housekeeping

This is a data summary document. I ran the files ``\stata_code\data_extraction_processing\extraction\commercial\00_cams_extraction.do`` and ``00_extraction_wrapper.do``

I also ran many of the  files in ``\stata_code\data_extraction_processing\analysis`` to do some exploratory analysis. 

Warnings: There is something funky going on with VA "status=PZERO" starting in 2021.  There is also something funky going on with Delaware landings, but there are no hullids/permit numbers so I don't think this will throw anything off at the vessel level.

--->

```{r global_options, include=FALSE}

 library("foreign")
 library("here")
 library("tidyverse")
 library("scales")
 library("knitr")
 library("lubridate")
 library("kableExtra")
 library("haven")
 library("mapview")
 library("sf")
here::i_am("writing/BSB_data_summary.Rmd")

#############################################################################
#knitr options

knitr::opts_chunk$set(echo=FALSE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, 
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)


#############################################################################
my_images<-here("images")

descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")
```

```{r readin_stat_areas, eval=FALSE}


# Read in statistical areas
stat_areas_location<-here("data_folder","external","shapefiles","Statistical_Areas_2010.shp")
stat_areas <- st_read(dsn = stat_areas_location)

#trim out some stat areas that I don't care about.
stat_areas<-stat_areas %>%
  dplyr::filter(Id>=460) %>% # Canada
   dplyr::filter(Id<=711) %>% # SERO
  dplyr::filter(!Id %in% c(640,650,660,670,680) ) #Offshore

```

# Summary: Research Question, Motivation, Extensions, and Potential Problems

## Research Question

How do the state-level landings restrictions impact the profits of firms that fish for this species and places that they land fish? What sort of gains might be realized if these restrictions were changed to allow fishing fleets more flexibility to follow the fish? Who would get those gains, and what sectors would be negatively impacted?

## Motivation

Since 2003, the commercial black sea bass quota has been allocated to individual states by the ASMFC.  Quota charges are based on the state in which the catch was landed and monitored  through the dealer reporting systems. States have chosen a variety of regulatory systems for their commercial fleets, including input controls (mesh or vent sizes), daily- or weekly- possession limits, output controls (individual-, seasonal- , and aggregate quotas), and minimum sizes.

All states have implemented limited access, which restricts both the ability of firms to fish in state waters and the ability of firms to land fish in their state, regardless of where fish were caught.  These restrictions are intended to preserve state-level quota for state vessels.  When stocks of fish move, these restrictions can create a geographic mismatch between the fishing fleet and the biomass, increasing costs (perhaps dramatically).

I don't think I can separate this from Summer Flounder and Scup state-level landing restrictions.

More broadly, this is a story about managing the right thing and the unintended consequences of managing the wrong thing.

# Methods, Identification Strategy, and Pitfalls

We will estimate a model of fishing location and landing choice for firms in the black sea bass fishery.  Location choice models will be fairly straightforward, but we will experiment with letting the expected catch from a part of the ocean depend on a "local" measure of biomass. Utility from landing locations will be modeled as a function of distances, prices, and ``home state.``  The home state variable will be coded as 1 if a vessel is registered in that state and 0 otherwise.  We will simulate out of sample at ``home state=1`` to understand how the distribution of landings might change if there were no landings restrictions. 

## Pitfalls

* We need enough observations in the dataset where a firm lands "out of state."  I have between 350 and 500 vessels annualy in my "black sea bass landings" dataset.  Approximately 1/4th to 1/3rd switch states at least 1 time in a year, which should be enough variation to estimate an econometric model. 
  * This dataset only includes trips that landed black sea bass, so I need to narrow down a "fleet" and then pull in that fleet's full set of trips. 
  * I've screened out some of the data yuck -- 
  ``permit in (190998,290998,390998,490998)``, ``hullid =0`` (and similar), and ``hullid" = FROM_SHORE`` 

  * there's no "pounds landed" filter here, I suspect alot of trips are landing a small amount of BSB.
* the ``homestate'' variable is very unstructural. 
  * Firms can, at some cost, get a permit from another state.  There's an up-front cost plus an ongoing cost to renew that around 500-1000/year.
  * These are 'short-run' *ceteris paribus* effects.  
  * Most state permits are issued to the person, not to the vessel. Could a permitted person get on an "out of state" vessel?
  * Can I model whether a firm chooses to get an "out-of-state" permit?
  * Is this a kind of latent class thing, since I might not observe it?
* The regulatory history is *really* complicated.
* Trips and catch by the non-federally permits is so tough. Can we get something about this?



# Related

* Landing location choices
* Fishing location choices
* Summer Flounder "climate informed" production function (Teng, Speir, Lipton)

# Deep Background 

<!--- Include the next 2 child documents.   --->
```{r, child=here("writing", c("BSB_history.Rmd", "BSB_economic_background.Rmd")),  eval=FALSE}
```



\newpage

# References
<div id="refs"></div>

