---
title: "Black Sea Bass: A Random Forest, read in tune results, finalfit, and validate"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: nocluster
    choices:
    - nocluster
    - TS_nocluster

---

# Summary and Housekeeping


 My "final fit" bombed out because I didn't have drive space.  This is code to
 
1. Read in the tune result and eyeball them.
2. Pick the best set of tuning parameters.
3. Fit a final model. 
4. Calibrate probabilities using a 'testing sample'
5. Validate on a validation sample. 

<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	Read in outputs of machine learning models 
* Inputs: These three can be adjusted by changing the "modeltype"
*   - Data file
*   - tuning results file
*   - final results file


* Outputs:
*   - Graphs of tuning quality and final model fit
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


**********************************************************************

 
 --->
```{r global_options, include=FALSE}

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/final_fit_and_validate.Rmd")


# You've estimated a few difference models, this sets the 
if (modeltype=="standard"){
  data_pattern<-"data_split"
  tuning_pattern<-"BSB_ranger_tune"
  final_pattern<-"BSB_ranger_results"
} else if (modeltype=="nocluster"){
  data_pattern<-"nocluster_data_split"
  tuning_pattern<-"BSB_ranger_nocluster_tune"
  final_pattern<-"BSB_ranger_nocluster_results"
} else if (modeltype=="TS_nocluster"){
  data_pattern<-"nocluster_TS_data_split"
  tuning_pattern<-"BSB_ranger_TS_nocluster_tune"
  final_pattern<-"BSB_ranger_TS_nocluster_results"
}else {
  stop("Unknown modeltype")
}





#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)

# My "final fit" bombed out because I didn't have drive space.  
# finalfit_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{final_pattern}*Rds")))
# finalfit_vintage<-gsub(final_pattern,"",finalfit_vintage)
# finalfit_vintage<-gsub(".Rds","",finalfit_vintage)
# finalfit_vintage<-max(finalfit_vintage)

finalfit_vintage<-tuning_vintage

vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)



```

The modeltype is `r params$modeltype`, which is hopefully the same as `r modeltype`.

Data vintage is `r data_vintage_string`.

Tuning vintage is `r tuning_vintage`.

Final fit vintage is `r finalfit_vintage`.

```{r load_in_data and tuning_results, include=FALSE}
# this was created with data_prep_ml.Rmd
data_split<-readr::read_rds(file=here("results","ranger",glue("{data_pattern}{data_vintage_string}.Rds")))
tune_res<-readr::read_rds(file=here("results","ranger",glue("{tuning_pattern}{tuning_vintage}.Rds")))
oos_data<-read_rds(file=here("data_folder","main","commercial",glue("BSB_unclassified_dataset{raw_oos_data_vintage_string}.Rds")))

train_data <- training(data_split)
test_data <- testing(data_split)
validation_data <- validation(data_split)

source(here("R_code","analysis","fit_random_forest","BSB.Classification.Recipe.R"))
search_type<-"Initial"


source(here("R_code","analysis","fit_random_forest","BSB.Workflow.Setup.R"))


```

```{r final_fit}

bayesian_results<-autoplot(tune_res, type = "performance") +
  labs(title = "Did Bayesian optimization converge?")

 ggsave(here("results","ranger","tune",glue("bayesian_tune_{modeltype}{tuning_vintage}.png")), plot=bayesian_results)

 write_rds(bayesian_results, file=here("results","ranger","final",glue("bayesian_tune_{modeltype}{finalfit_vintage}.Rds")))


 #It didn't find anything better.

# Select the best Rforest based on log loss from the 10 folds.  Do a final fit on the full training dataset, predict on the validation dataset. Save the data

best_tree <- tune_res %>%
  select_best(metric = "mn_log_loss")

best_config<-best_tree %>%
  pull(.config)

best_config
```

```{r }

# eyeball the metrics 
all.metrics<-tune_res %>%
  collect_metrics() 

all.metrics %>% filter(.metric=="mn_log_loss") %>% arrange(mean) %>% slice_head(n=15)
```
The ``.iter`` value is the iteration log from the bayesian search. ``iter==0`` means that that particular hyperparamter configuration was part of the initial grid search. The 1st Bayesian search is really keyed on small min_n, while the grid search really likes a higher one.  

For noclusterTS, the grid search was pretty good, with mtry=18, min_n=17 producing a mn_log_loss of 0.391.  The bayesian search found 26,15, 0.388 on iteration 8 and then failed to improve on that for 10 iterations, so  it stopped.


```{r finalize_model, eval=FALSE}
# finalize model by picking the best model hyperparameters
# I have an 'if' exists statement going on here, because it takes a little while to fit the model. If I've already fit the model, I'll just read it in.
# the first time, my code bombed out because I didn't have storage space. Now, the full fit worked, so it's safer to eval=FALSE this chunk

final_wf <- 
  BSB.Ranger.Workflow %>% 
  finalize_workflow(best_tree)

if (file.exists(here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))==FALSE){
# Final model fitting on the full training dataset 
final_fit <- 
  final_wf %>%
  last_fit(data_split, metrics=class_and_probs_metrics) 


write_rds(final_fit, file=here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))
}
```

```{r brier_class, eval=FALSE}

# Brier class is broken for me, but being fixed by the yardstick devs
# Now that I'm tuning over a grid, this makes less sense.
best_tree <- tune_res %>%
  select_best(metric = "brier_class")

opt_mtry<-best_tree$mtry

brier<-all.metrics %>%
  filter(.metric == "brier_class") %>%
  select(mean, mtry) %>%
  rename(brier_class=mean) 

 brier %>% ggplot(aes(mtry, brier_class)) +
  geom_point(show.legend = TRUE) +
   geom_text_repel(aes(label = round(brier_class,4)), size=3) 


# ggsave(here("results","ranger","tune",glue("brier_class{modeltype}{tuning_vintage}.png")), plot=last_plot())


 
  
brier<-tune_res %>%
  collect_metrics(summarize=FALSE) %>%
  filter(.metric == "brier_class") 

brier %>% ggplot(aes(x=as.factor(mtry), y=.estimate)) +
  geom_boxplot(show.legend = TRUE) 

#  ggsave(here("results","ranger","tune",glue("brier_class_indivd_{modeltype}{tuning_vintage}.png")), plot=last_plot())

 
```

Multinomial log-loss. Since I'm tuning over mtry and min_n, the 1d plot doesn't make as much sense. It's better to look at the table. 

```{r log_loss}
mn_log_loss<-all.metrics %>%
  filter(.metric == "mn_log_loss") %>%
#  filter(mtry <=10) %>%
  select(mean, mtry, min_n) %>%
  rename(mn_log_loss=mean)

log_loss<- mn_log_loss %>% ggplot(aes(mtry, mn_log_loss)) +
  geom_point(show.legend = TRUE) + 
    geom_text_repel(aes(label = round(mn_log_loss,4)), size=3) 


 ggsave(here("results","ranger","tune",glue("mn_log_loss_mtry_{modeltype}{tuning_vintage}.png")), plot=log_loss)

 write_rds(log_loss, file=here("results","ranger","final",glue("mn_log_loss_mtry_{modeltype}{finalfit_vintage}.Rds")))

 
min_n<-  mn_log_loss %>% ggplot(aes(min_n, mn_log_loss)) +
  geom_point(show.legend = TRUE) + 
    geom_text_repel(aes(label = round(mn_log_loss,4)), size=3) 

  ggsave(here("results","ranger","tune",glue("mn_log_loss_min_n_{modeltype}{tuning_vintage}.png")), plot=min_n)
 write_rds(min_n, file=here("results","ranger","final",glue("mn_log_loss_min_n_{modeltype}{finalfit_vintage}.Rds")))

 p<- plot_ly(mn_log_loss, 
        x = ~mtry, 
        y = ~min_n, 
        z = ~mn_log_loss,
        type = "mesh3d", 
        intensity=~mn_log_loss,
        colorscale="Hot",
        reversescale=TRUE
)

 saveWidget(p, here("results","ranger","tune",glue("mn_log_loss{modeltype}{tuning_vintage}.html")), selfcontained = TRUE)

  
```

The initial tune of TS nocluster showed it was pretty insensitive to min_n.
Same thing for the nocluster. For both, i feel comfortable adjusting the mtry grid to 5-35.

I need to expand the min.n search though.


Are there any warnings from the tuning?

```{r warnings}
extracts<-tune_res %>%
  collect_notes()
extracts
not_good<-unique(extracts$id)
not_good
```

# Some in-sample results from Tuning

```{r roc}
# A ROC curve. I'm not sure if this is quite right, but I think it is.
predictions<-tune_res %>%
  collect_predictions()

# pull the probability names
prob_names<-colnames(predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


roc_plot<-predictions %>%
  dplyr::filter(.config==best_config) %>%
  group_by(id) %>%
  #dplyr::filter(!id %in% not_good) %>%
  roc_curve(market_desc, any_of(prob_names)) %>%
  ggplot(aes(x=1-specificity,y=sensitivity, colour=id)) +
   facet_wrap(~.level) + 
  geom_line(show.legend = TRUE) 

print(last_plot())

ggsave(here("results","ranger","tune",glue("roc_tune_{modeltype}{tuning_vintage}.png")), plot=roc_plot)


pr_plot<-predictions %>%
  group_by(id) %>%
  dplyr::filter(.config==best_config) %>%
#  dplyr::filter(!id %in% not_good) %>%
  pr_curve(market_desc, any_of(prob_names))%>% ggplot(aes(x=recall,y=precision, colour=id)) +
   facet_wrap(~.level) + 
  geom_line(show.legend = TRUE) 

ggsave(here("results","ranger","tune",glue("pr_curve_tune_{modeltype}{tuning_vintage}.png")), plot=pr_plot)


  
```

# Load in the final model

read in the final model, extract the recipe and the predictors, extract the predictions

```{r load_final_model}

rm(tune_res)

final_fit<-read_rds(file=here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))


# What is the workflow
final_workflow<-extract_workflow(final_fit)

# And the recipe
extract_recipe(final_fit)

preds<-final_fit$.workflow[[1]]$pre$actions$recipe$recipe$var_info %>%
 dplyr::filter(role=="predictor")

num_preds<-nrow(preds)

final_predictions<-final_fit %>%
  collect_predictions()

# pull the probability names
prob_names<-colnames(predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


# look at the fit metrics
final_fit_metrics<-final_fit %>%
  collect_metrics()
final_fit_metrics

all.metrics %>%
  dplyr::filter(.config==best_config) 

```

Log loss of the final fit is slightly better than that of the average of the individual folds. This is normal/reasonable, since it uses more data than each of the individual folds.  It's not so different that alarm bells go off.


## Validation Diagnostics

Look at the VIP, ROC, and PR curves

```{r VIP}
# Look at variable importance. I want to see all the predictors.
# The scale is different from previous versions, because I switched the variable importance method in the Workflow/recipe to "Permutation"
vip_plot<-final_fit %>% 
  extract_fit_parsnip() %>% 
  vip(num_features = num_preds)
ggsave(here("results","ranger","final",glue("vip{modeltype}{finalfit_vintage}.png")), plot=vip_plot)


```
Variable importance measures are the Prices and the Dealer proporption variables. 


```{r ROC_Curve}

# Reciever operator curve (True positive vs False Positive) for the final model evaluated on the validation data
final_roc<-final_predictions %>%
  roc_curve(market_desc, any_of(prob_names))  %>%
  autoplot()
ggsave(here("results","ranger","final",glue("roc_{modeltype}{finalfit_vintage}.png")), plot=final_roc)
```
```{r PR_Curve}
# PR curve
final_pr<-final_predictions %>%
  pr_curve(market_desc,  any_of(prob_names)) %>%
  autoplot()
ggsave(here("results","ranger","final",glue("pr_curve{modeltype}{finalfit_vintage}.png")), plot=final_pr)
```

# Validation and Calibration 

Use the Validation set to (1 assess how well the model fits) and 2. Calibrate if I have systematic under or over prediction.


```{r Validation_predictions}
# Apply the Final fit to the validation dataset

validation_data <- augment(final_workflow,validation_data)

#The validation plots indicate characteristic issues with probability trees. All models are underconfident. Groups of observations with a 75% predicted probability of Jumbo have a true probability of Jumbo of almost 90%.  Groups of observations with a predicted probability of ~15% probabilty of Jumbo have a true probability closer to 10%.

# Calibration should fix this.

cal_window<-validation_data %>%
  cal_plot_windowed(truth = market_desc, step_size = 0.025, include_points = FALSE)
cal_window
ggsave(here("results","ranger","final",glue("cal_windowed_{modeltype}{finalfit_vintage}.png")), plot=cal_window)

cal_breaks<-validation_data %>%
  cal_plot_breaks(truth = market_desc, step_size = 0.025, include_points = FALSE)
ggsave(here("results","ranger","final",glue("cal_breaks_{modeltype}{finalfit_vintage}.png")), plot=cal_breaks)

cal_logistic<-validation_data %>%
  cal_plot_logistic(truth = market_desc)
cal_logistic
ggsave(here("results","ranger","final",glue("cal_logistic_{modeltype}{finalfit_vintage}.png")), plot=cal_logistic)

```


Experiment with different methods of calibration. Beta fits a 3 parameter beta model (2 shape, 1 location), performing 4 distinct 1 vs all models and then re-normalalizing the results.  Iso fits an isotonic regression (4 separate 1 vs all again).  cal_estimate_multinom fits logistic splines and seemed to work better than the version that just fit a multinomial logistic.

```{r calibration_models}

calibrate_beta <- cal_estimate_beta(validation_data, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)

calibrate_iso <- cal_estimate_isotonic(validation_data, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)

calibrate_multinom <- cal_estimate_multinomial(validation_data, 
                                        truth=market_desc, 
                                        estimate=.pred_Jumbo:.pred_Small,
                                        smooth=TRUE)

# weights <- validation_data$lndlb
# 
# calibrate_multinomW <- cal_estimate_multinomial(validation_data, 
#                                         truth=market_desc, 
#                                         estimate=.pred_Jumbo:.pred_Small,
#                                         smooth=FALSE, parameters=NULL, weights=lndlb)


```

```{r explore_validations, eval=FALSE}
# This section 'unweights' the validation data and then tries to fit a calibration.
# It doesn't work too well, in part because the initial cal_plot is non-monotone. 
# The observation-level model is textbook smooth, the unweighted one is not. 
# This makes the parametric (beta, multi SMOOTH=FALSE) fit poorly
# The multi Smooth and isotonic "should" fit better. 
# The isotonic takes forever to run, because it has to do alot of comparisons.

validation_expanded <- validation_data %>%
  uncount(weights = weighting)

# this dataset is massive. take a 5% sample, just to see what is going on.

validation_sub<-initial_split(validation_expanded,prop=0.90)
tt<-testing(validation_sub)


tt %>%
    cal_plot_windowed(truth = market_desc,step_size = 0.05)

tt %>%
  cal_plot_breaks(truth = market_desc, step_size = 0.025)


calibrate_betaTT <- cal_estimate_beta(tt, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)

calibrate_multiTT <- cal_estimate_multinomial(tt, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=TRUE)

calibrate_isoTT <- cal_estimate_isotonic(tt, 
                                   truth=market_desc, 
                                   estimate=.pred_Jumbo:.pred_Small,
                                   smooth=FALSE)


WT_data_calib_applied<-  tt %>%
  cal_apply(calibrate_multiTT)


WT_data_calib_applied %>%
    cal_plot_windowed(truth = market_desc,step_size = 0.05)



W_ESPR_raw<-tt %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

W_ESPR_multi<-WT_data_calib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)


CalLoss<-W_ESPR_raw-W_ESPR_multi
rCalLoss<-100*CalLoss/W_ESPR_raw

CalLoss
rCalLoss


# calibrate_iso <- cal_estimate_isotonic(validation_expanded, 
#                                    truth=market_desc, 
#                                    estimate=.pred_Jumbo:.pred_Small,
#                                    smooth=FALSE)





validation_exp_data_isocalib_applied <-
  validation_expanded %>%
  cal_apply(calibrate_iso)

validation_exp_data_isocalib_applied %>%
  cal_plot_breaks(truth = market_desc, step_size = 0.025)


```





```{r apply_calibration}
# This is fine for testing out things, but the final judging is done on the 'testing' dataset

validation_data_isocalib_applied <-
  validation_data %>%
  cal_apply(calibrate_iso)

validation_data_betacalib_applied <-
  validation_data %>%
  cal_apply(calibrate_beta)

validation_data_multicalib_applied <-
  validation_data %>%
  cal_apply(calibrate_multinom)


iso_applied_window<-cal_plot_windowed(validation_data_isocalib_applied,
                  truth = market_desc, step_size = 0.05, include_points =FALSE)
iso_applied_window

ggsave(here("results","ranger","final",glue("cal_isovalidate_window{modeltype}{finalfit_vintage}.png")), plot=iso_applied_window)

write_rds(iso_applied_window, file=here("results","ranger","final",glue("cal_isovalidate_window{modeltype}{finalfit_vintage}.Rds")))


#Multi looks a little rough
multi_applied_window<-cal_plot_windowed(validation_data_multicalib_applied,
                  truth = market_desc, step_size = 0.05, include_points =FALSE)
multi_applied_window
ggsave(here("results","ranger","final",glue("cal_multicalibrate_window{modeltype}{finalfit_vintage}.png")), plot=multi_applied_window)

write_rds(multi_applied_window, file=here("results","ranger","final",glue("cal_multicalibrate_window{modeltype}{finalfit_vintage}.Rds")))


beta_applied_window<-cal_plot_windowed(validation_data_betacalib_applied,
                  truth = market_desc, step_size = 0.05, include_points =FALSE)
ggsave(here("results","ranger","final",glue("cal_betavalidate_window{modeltype}{finalfit_vintage}.png")), plot=beta_applied_window)

write_rds(beta_applied_window, file=here("results","ranger","final",glue("cal_betavalidate_window{modeltype}{finalfit_vintage}.Rds")))


# Compute the Calibration Loss (from Ferrer and Ramos "Evaluating Posterior Probabilities")
# (ESPR_raw - ESPR_cal)/ESPR_raw

ESPR_raw<-validation_data %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_caliso<-validation_data_isocalib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_calbeta<-validation_data_betacalib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_calmulti<-validation_data_multicalib_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

# the difference between these metrics is the calibration loss (A little weird). But basically, the "goodness of fit metric" improves when I apply a calibration. the multinomial model doesn't work so well. But Iso and Beta do. Iso is slightly better, so we will use that.
ESPR_raw
ESPR_caliso
ESPR_calbeta
ESPR_calmulti

CalLoss<-ESPR_raw-ESPR_caliso
rCalLoss<-100*CalLoss/ESPR_raw

CalLoss
rCalLoss
```




```{r final_model_testing}


test_data2 <- augment(final_workflow,test_data )





#Plot the "raw" predictions
uncalibrated<-cal_plot_windowed(test_data2, truth = market_desc, step_size = 0.05,include_points=FALSE)

ggsave(here("results","ranger","final",glue("validation_raw_{modeltype}{finalfit_vintage}.png")), plot=uncalibrated)
write_rds(uncalibrated, file=here("results","ranger","final",glue("validation_raw_{modeltype}{finalfit_vintage}.Rds")))


test_data_calibration_applied <-
  test_data2 %>%
  cal_apply(calibrate_iso)

# Plot the calibrated predictions

calibrated<-cal_plot_windowed(test_data_calibration_applied, truth = market_desc, step_size = 0.05,include_points=FALSE)

calibrated

ggsave(here("results","ranger","final",glue("validation_calibrated_{modeltype}{finalfit_vintage}.png")), plot=calibrated)

write_rds(calibrated, file=here("results","ranger","final",glue("validation_calibrated_{modeltype}{finalfit_vintage}.Rds")))



#cal_plot_breaks(test_data_calibration_applied, truth = market_desc, num_breaks=20,include_points=FALSE)

# Compute the Calibration Loss (from Ferrer and Ramos "Evaluating Posterior Probabilities")
# (ESPR_raw - ESPR_cal)/ESPR_raw

ESPR_raw<-test_data2 %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)


ESPR_cal<-test_data_calibration_applied %>%
  mn_log_loss(market_desc,.pred_Jumbo:.pred_Small) %>%
  pull(.estimate)

ESPR_raw
ESPR_cal

CalLoss<-ESPR_raw-ESPR_cal
rCalLoss<-100*CalLoss/ESPR_raw

#  The iso calibration improves model fit a bit. 

CalLoss
rCalLoss

```

# Calibrated Predictions

In this section, I predict out of sample on my 'test' dataset with the calibration applied 


```{r extract_predictions}
### pull a few categories, compute a prob*lndlb variable, sum 


# keep just a few columns
test_data_calibration_applied<-test_data_calibration_applied %>%
  select(c( weighting, market_desc, lndlb, stockarea, year, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the landed pounds to get predicted pounds in each class. 
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    test_data_calibration_applied[[paste0("pred_", l)]] <- test_data_calibration_applied[[paste0(".pred_", l)]] * test_data_calibration_applied$lndlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}
# Sum up the predictions by year, market category, and stockarea. 
test_predictions<-test_data_calibration_applied %>%
   group_by(stockarea, year,market_desc) %>%
  select(-.pred_class) %>%
  mutate(transactions=1) %>%
  summarise(across(c(starts_with(c("pred_",".pred_")), "lndlb", "transactions"), \(x) sum(x, na.rm = TRUE))) %>%
   ungroup()
test_predictions$modeltype<-modeltype

test_predictions<-test_predictions %>%
  mutate(market_desc=fct_relevel(market_desc,"Jumbo","Large","Medium","Small")) 

write_rds(test_predictions, file=here("results","ranger",glue("aggregate_test_predictions_{modeltype}{finalfit_vintage}.Rds")))

#test_predictions<-test_predictions %>%
#  dplyr::filter(year %in% c(2021,2022,2023,2024))

```

## Probabalistic Predictions of the class of each transaction

Here are the predictions (transactions and mt) for the hold-out sample, using the probability calibrations.

```{r topline_validation}

aggregate_transactions<-test_predictions %>%
    summarise(across(c(starts_with(".pred_")), sum)) %>%
  pivot_longer(cols=starts_with(".pred_"), names_to="market_desc", names_prefix=".pred_",values_to="ObsPredicted")


aggregate_test_predictions<-test_predictions %>%
    summarise(across(c(starts_with("pred_")), sum)) %>%
  pivot_longer(cols=starts_with("pred_"), names_to="market_desc", names_prefix="pred_",values_to="predicted")


true<-test_predictions %>%
    group_by(market_desc) %>%
    summarise(lndlb=sum(lndlb),
              transactions=sum(transactions))

Testmkt_preds<-aggregate_transactions %>%
  left_join(true, by=join_by(market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  left_join(aggregate_test_predictions, by=join_by(market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted,
         predicted_transactions=ObsPredicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01),
         `Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )


transaction_predictions<-Testmkt_preds %>%
  select(market_desc,transactions, predicted_transactions, `Transaction error (%)`)


knitr::kable(transaction_predictions, caption='Calibrated Predictions on the 15% Validation Sample (transaction count).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```

```{r weighted_predictions}
weighted_predictions<-Testmkt_preds %>%
  select(market_desc,true_mt, predicted_mt, `mt error (%)`)

knitr::kable(weighted_predictions, caption='Calibrated Predictions on the 15% Validation Sample (mt).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```


Here are the predictions (transactions and mt) for the hold-out sample, without the probability calibrations.

```{r aggregate_test2_predictions}
# Subset data, construct predicted pounds, and aggregate


### pull a few categories, compute a prob*lndlb variable, sum 


# keep just a few columns
test_data_nocalibration_applied<-test_data2%>%
  select(c( weighting, market_desc, lndlb, stockarea, year, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the landed pounds to get predicted pounds in each class. 
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    test_data_nocalibration_applied[[paste0("pred_", l)]] <- test_data_nocalibration_applied[[paste0(".pred_", l)]] * test_data_nocalibration_applied$lndlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}
# Sum up the predictions by year, market category, and stockarea. 
nocal_predictions<-test_data_nocalibration_applied%>%
   group_by(stockarea, year,market_desc) %>%
  select(-.pred_class) %>%
  mutate(transactions=1) %>%
  summarise(across(c(starts_with(c("pred_",".pred_")), "lndlb", "transactions"), \(x) sum(x, na.rm = TRUE))) %>%
   ungroup()
nocal_predictions$modeltype<-modeltype

nocal_predictions<-nocal_predictions%>%
  mutate(market_desc=fct_relevel(market_desc,"Jumbo","Large","Medium","Small")) 


```







```{r no_cal_validations}

aggregate_nocal_transactions<-nocal_predictions%>%
    summarise(across(c(starts_with(".pred_")), sum)) %>%
  pivot_longer(cols=starts_with(".pred_"), names_to="market_desc", names_prefix=".pred_",values_to="ObsPredicted")


aggregate_nocal_weight_predictions<-nocal_predictions %>%
    summarise(across(c(starts_with("pred_")), sum)) %>%
  pivot_longer(cols=starts_with("pred_"), names_to="market_desc", names_prefix="pred_",values_to="predicted")


true<-test_data2%>%
    mutate(transactions=1) %>%
    group_by(market_desc) %>%
    summarise(lndlb=sum(lndlb),
              transactions=sum(transactions))

nocalibrate_results<-aggregate_nocal_weight_predictions%>%
  left_join(true, by=join_by(market_desc==market_desc))

nocalibrate_results<-nocalibrate_results%>%
  left_join(aggregate_nocal_transactions, by=join_by(market_desc==market_desc))

nocalibrate_results<-nocalibrate_results%>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted,
         predicted_transactions=ObsPredicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01),
         `Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )
  

transactions_nocalibrate_predictions<-nocalibrate_results %>%
  select(market_desc,transactions, predicted_transactions, `Transaction error (%)`)


knitr::kable(transactions_nocalibrate_predictions, caption='UnCalibrated Predictions on the 15% Validation Sample (transactions).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```


```{r weighted_no_calibrate_predictions}
weights_nocalibrate_predictions<-nocalibrate_results %>%
  select(market_desc,true_mt, predicted_mt, `mt error (%)`)
knitr::kable(weights_nocalibrate_predictions, caption='UnCalibrated Predictions on the 15% Validation Sample (transactions).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```









## Lets take a look at the validation predictions by stockarea



```{r Validation_Weight_Predictions_by_stockarea}

# Aggregate to the stockarea- market category

aggregate_test_predictions2<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(across(c(starts_with(".pred_"),lndlb), sum)) %>%
  arrange(stockarea,market_desc)

# true landings by stockarea and market_category
trueTest<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(lndlb=sum(lndlb))


# aggregate predictions, in pounds, by stockarea and market_category
Tp1<-test_predictions %>%
  group_by(stockarea) %>%
    summarise(across(c(starts_with("pred_")), sum))

Tp2<-Tp1  %>%
  pivot_longer(cols=!stockarea, names_to="market_desc", names_prefix="pred_",values_to="predicted")


Testmkt_preds<-Tp2 %>%
  left_join(trueTest, by=join_by(stockarea==stockarea, market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01) )


knitr::kable(Testmkt_preds, caption='Calibrated Predictions on the 15% Validation Sample (mt). Disaggregated to Stock area',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```




```{r Validation_Transaction_Predictions_by_stockarea}


# true landings by stockarea and market_category
trueTest<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(transactions=sum(transactions))


# aggregate predictions, in transactions, by stockarea and market_category
Tp1<-test_predictions %>%
  group_by(stockarea) %>%
    summarise(across(c(starts_with(".pred_")), sum))

Tp2<-Tp1  %>%
  pivot_longer(cols=!stockarea, names_to="market_desc", names_prefix=".pred_",values_to="predicted")


Testmkt_preds<-Tp2 %>%
  left_join(trueTest, by=join_by(stockarea==stockarea, market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  rename(predicted_transactions=predicted) %>%
  mutate(`Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )


knitr::kable(Testmkt_preds, caption='Calibrated Predictions on the 15% Validation Sample (mt). Disaggregated to Stock area',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  
```






<!---
\newpage
--->
# References
<div id="refs"></div>





# Appendix{-}