---
title: "Black Sea Bass: A Random Forest, read in tune results and perform the final fit"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: nocluster
    choices:
    - nocluster
    - TS_nocluster
---

# Summary and Housekeeping


 My "final fit" bombed out because I didn't have drive space.  This is code to
 
1. Read in the tune result and eyeball them.
2. Pick the best set of tuning parameters.
3. Fit a final model. 
4. make some figures.


<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	Read in outputs of machine learning models 
* Inputs: These three can be adjusted by changing the "modeltype"
*   - Data file
*   - tuning results file
*   - final results file


* Outputs:
*   - Graphs of tuning quality and final model fit
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


**********************************************************************

 
 --->
```{r global_options, include=FALSE}

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/final_fit.Rmd")


# You've estimated a few difference models, this sets the 
if (modeltype=="standard"){
  data_pattern<-"data_split"
  tuning_pattern<-"BSB_ranger_tune"
  final_pattern<-"BSB_ranger_results"
} else if (modeltype=="nocluster"){
  data_pattern<-"nocluster_data_split"
  tuning_pattern<-"BSB_ranger_nocluster_tune"
  final_pattern<-"BSB_ranger_nocluster_results"
} else if (modeltype=="TS_nocluster"){
  data_pattern<-"nocluster_TS_data_split"
  tuning_pattern<-"BSB_ranger_TS_nocluster_tune"
  final_pattern<-"BSB_ranger_TS_nocluster_results"
}else {
  stop("Unknown modeltype")
}





#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)

# My "final fit" bombed out because I didn't have drive space.  
# finalfit_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{final_pattern}*Rds")))
# finalfit_vintage<-gsub(final_pattern,"",finalfit_vintage)
# finalfit_vintage<-gsub(".Rds","",finalfit_vintage)
# finalfit_vintage<-max(finalfit_vintage)

finalfit_vintage<-tuning_vintage

vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)



```

The modeltype is `r params$modeltype`, which is hopefully the same as `r modeltype`.

Data vintage is `r data_vintage_string`.

Tuning vintage is `r tuning_vintage`.

Final fit vintage is `r finalfit_vintage`.

```{r load_in_tuning_results, include=FALSE}
tune_res<-readr::read_rds(file=here("results","ranger",glue("{tuning_pattern}{tuning_vintage}.Rds")))
#oos_data<-read_rds(file=here("data_folder","main","commercial",glue("BSB_unclassified_dataset{raw_oos_data_vintage_string}.Rds")))

#source(here("R_code","analysis","fit_random_forest","BSB.Classification.Recipe.R"))
#search_type<-"Initial"


#source(here("R_code","analysis","fit_random_forest","BSB.Workflow.Setup.R"))


```


# Tuning

How well did the tuning work?




```{r check_convergence}
# Did the bayesian model converge or is one of the grid-searchs best?

bayesian_results<-autoplot(tune_res, type = "performance") +
  labs(title = "Did Bayesian optimization converge?")

 ggsave(here("results","ranger","tune",glue("bayesian_tune_{modeltype}{tuning_vintage}.png")), plot=bayesian_results)

rm(bayesian_results)

 #It didn't find anything better.

# Select the best Rforest based on log loss from the 10 folds.  Do a final fit on the full training dataset, predict on the validation dataset. Save the data

best_tree <- tune_res %>%
  select_best(metric = "mn_log_loss")

best_config<-best_tree %>%
  pull(.config)

best_config
```

# do the metrics look okay?

```{r }

# eyeball the metrics 
all.metrics<-tune_res %>%
  collect_metrics() 

all.metrics %>% filter(.metric=="mn_log_loss") %>% arrange(mean) %>% slice_head(n=15)
```
The ``.iter`` value is the iteration log from the bayesian search. ``iter==0`` means that that particular hyperparamter configuration was part of the initial grid search. The 1st Bayesian search is really keyed on small min_n, while the grid search really likes a higher one.  

For noclusterTS, the grid search was pretty good, with mtry=18, min_n=17 producing a mn_log_loss of 0.391.  The bayesian search found 26,15, 0.388 on iteration 8 and then failed to improve on that for 10 iterations, so  it stopped.


```{r finalize_model, eval=FALSE}
# finalize model by picking the best model hyperparameters
# I have an 'if' exists statement going on here, because it takes a little while to fit the model. If I've already fit the model, I'll just read it in.
# the first time, my code bombed out because I didn't have storage space. Now, the full fit worked, so it's safer to eval=FALSE this chunk


source(here("R_code","analysis","fit_random_forest","BSB.Classification.Recipe.R"))
search_type<-"Initial"
source(here("R_code","analysis","fit_random_forest","BSB.Workflow.Setup.R"))

final_wf <- 
  BSB.Ranger.Workflow %>% 
  finalize_workflow(best_tree)

if (file.exists(here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))==FALSE){
# Final model fitting on the full training dataset 
final_fit <- 
  final_wf %>%
  last_fit(data_split, metrics=class_and_probs_metrics) 


}
gc()
```

```{r brier_class, eval=FALSE}

# Brier class is broken for me, but being fixed by the yardstick devs
# Now that I'm tuning over a grid, this makes less sense.
best_tree <- tune_res %>%
  select_best(metric = "brier_class")

opt_mtry<-best_tree$mtry

brier<-all.metrics %>%
  filter(.metric == "brier_class") %>%
  select(mean, mtry) %>%
  rename(brier_class=mean) 

 brier %>% ggplot(aes(mtry, brier_class)) +
  geom_point(show.legend = TRUE) +
   geom_text_repel(aes(label = round(brier_class,4)), size=3) 


# ggsave(here("results","ranger","tune",glue("brier_class{modeltype}{tuning_vintage}.png")), plot=last_plot())


 
  
brier<-tune_res %>%
  collect_metrics(summarize=FALSE) %>%
  filter(.metric == "brier_class") 

brier %>% ggplot(aes(x=as.factor(mtry), y=.estimate)) +
  geom_boxplot(show.legend = TRUE) 

#  ggsave(here("results","ranger","tune",glue("brier_class_indivd_{modeltype}{tuning_vintage}.png")), plot=last_plot())

 
```

Multinomial log-loss. Since I'm tuning over mtry and min_n, the 1d plot doesn't make as much sense. It's better to look at the table. 

```{r log_loss}
mn_log_loss<-all.metrics %>%
  filter(.metric == "mn_log_loss") %>%
#  filter(mtry <=10) %>%
  select(mean, mtry, min_n) %>%
  rename(mn_log_loss=mean)

log_loss<- mn_log_loss %>% ggplot(aes(mtry, mn_log_loss)) +
  geom_point(show.legend = TRUE) + 
    geom_text_repel(aes(label = round(mn_log_loss,4)), size=3) 


 ggsave(here("results","ranger","tune",glue("mn_log_loss_mtry_{modeltype}{tuning_vintage}.png")), plot=log_loss)

rm(log_loss)
 
min_n<-  mn_log_loss %>% ggplot(aes(min_n, mn_log_loss)) +
  geom_point(show.legend = TRUE) + 
    geom_text_repel(aes(label = round(mn_log_loss,4)), size=3) 

  ggsave(here("results","ranger","tune",glue("mn_log_loss_min_n_{modeltype}{tuning_vintage}.png")), plot=min_n)
rm(min_n)
 p<- plot_ly(mn_log_loss, 
        x = ~mtry, 
        y = ~min_n, 
        z = ~mn_log_loss,
        type = "mesh3d", 
        intensity=~mn_log_loss,
        colorscale="Hot",
        reversescale=TRUE
)

 saveWidget(p, here("results","ranger","tune",glue("mn_log_loss{modeltype}{tuning_vintage}.html")), selfcontained = TRUE)
rm(p)
  
```

The initial tune of TS nocluster showed it was pretty insensitive to min_n.
Same thing for the nocluster. For both, i feel comfortable adjusting the mtry grid to 5-35.

I need to expand the min.n search though.


Are there any warnings from the tuning?

```{r warnings}
extracts<-tune_res %>%
  collect_notes()
extracts
not_good<-unique(extracts$id)
not_good
```

# Some in-sample results from Tuning

```{r roc}
predictions<-tune_res %>%
  collect_predictions()
rm(tune_res)

# pull the probability names
prob_names<-colnames(predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


roc_data<-predictions %>%
  dplyr::filter(.config==best_config) %>%
  group_by(id) %>%
  #dplyr::filter(!id %in% not_good) %>%
  roc_curve(truth=market_desc, 
            any_of(prob_names),
            case_weights=.case_weights) 

# --- 3. Colour palette — four classes ---
class_colours <- c(
  "Jumbo"  = "#1B6CA8",   # deep blue
  "Large"  = "#E05C2A",   # burnt orange
  "Medium" = "#2E8B57",   # sea green
  "Small"  = "#7B3F9E"    # purple
)

roc_mean <- roc_data %>%
  group_by(.level, specificity) %>%
  summarise(sensitivity = mean(sensitivity), .groups = "drop")


# --- 4. Build the figures ---
p_facet <- ggplot(roc_data,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level, group = interaction(.level, id)),
            linewidth = 0.3, alpha = 0.35, show.legend = FALSE) +
    geom_path(data = roc_mean,
            aes(colour = .level),
            linewidth = 1.0, alpha = 1.0, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
# --- 5. Save at ICES JMS specifications ---
ggsave(here("results","ranger","tune",glue("roc_tune_{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facet,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)



roc_dataUW<-predictions %>%
  dplyr::filter(.config==best_config) %>%
  group_by(id) %>%
  #dplyr::filter(!id %in% not_good) %>%
  roc_curve(truth=market_desc, 
            any_of(prob_names)) 
roc_dataUW <- roc_data %>%
  group_by(.level, specificity) %>%
  summarise(sensitivity = mean(sensitivity), .groups = "drop")



# --- 3. Colour palette — four classes ---
class_colours <- c(
  "Jumbo"  = "#1B6CA8",   # deep blue
  "Large"  = "#E05C2A",   # burnt orange
  "Medium" = "#2E8B57",   # sea green
  "Small"  = "#7B3F9E"    # purple
)

# --- 4. Build the figures ---
p_facet <- ggplot(roc_dataUW,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level), linewidth = 0.8, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
# --- 5. Save at ICES JMS specifications ---
ggsave(here("results","ranger","tune",glue("roc_tuneUW_{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facet,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)








pr_plot<-predictions %>%
  group_by(id) %>%
  dplyr::filter(.config==best_config) %>%
#  dplyr::filter(!id %in% not_good) %>%
  pr_curve(truth=market_desc, any_of(prob_names),
                       case_weights=.case_weights)  %>%
 ggplot(aes(x=recall,y=precision, colour=id)) +
   facet_wrap(~.level) + 
  geom_line(show.legend = TRUE) 

ggsave(here("results","ranger","tune",glue("pr_curve_tune_{modeltype}{tuning_vintage}.png")), plot=pr_plot)

rm(pr_plot)
  rm(predictions)
gc()
```
