---
title: "Black Sea Bass: Out of Sample Predictions after model fit"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: nocluster
    choices:
    - nocluster
    - TS_nocluster
---

# Summary and Housekeeping


Calibration and Validation

1. Read in the finalized model 
2. Calibrate probabilities using a 'testing sample'
3. Validate on a validation sample. 

<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	Read in outputs of machine learning models 
* Inputs: These three can be adjusted by changing the "modeltype"
*   - Data file
*   - tuning results file
*   - final results file


* Outputs:
*   - Graphs of tuning quality and final model fit
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


The RF model was fit with "estimate_randomforest_nocluster.R", which has a few dependencies, mostly for the recipe and workflow.
The model is finalized with final_fit and calibrated with "calibration and validation.Rmd"



**********************************************************************

 
 --->
```{r global_options, include=FALSE}

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())
conflicts_prefer(vip::vi)

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/out_of_sample_predictions.Rmd")

#handle model patterns
source(here("R_code","analysis","helpers","modeltype_patterns.R"))





#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)

# My "final fit" bombed out because I didn't have drive space.  
# finalfit_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{final_pattern}*Rds")))
# finalfit_vintage<-gsub(final_pattern,"",finalfit_vintage)
# finalfit_vintage<-gsub(".Rds","",finalfit_vintage)
# finalfit_vintage<-max(finalfit_vintage)

finalfit_vintage<-tuning_vintage

vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

vintage_string<-gsub("BSB_estimation_dataset","",vintage_string)
vintage_string<-gsub(".Rds","",vintage_string)
vintage_string<-max(vintage_string)


raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)

estimation_dataset<-readr::read_rds(file=here("data_folder","main","commercial",glue("BSB_estimation_dataset{vintage_string}.Rds")))


```


# Get Ready

read in the final model, extract the recipe and the predictors. Read in the probabilty calibration model. Read in the out-of-sample dataset and partition it based on the ``mark_in`` column. ``mark_in`` was used to mark the "classified" data as in- or out- of the estimation sample. Observations with odd prices (outlier high/low), outlier states (FL, CN, PA, SC) wer excluded. Observations from dealers that reported minimal variability in prices paid were also excluded.

```{r load_final_model}


final_fit<-read_rds(file=here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))


# What is the workflow
final_workflow<-extract_workflow(final_fit)

# And the recipe
extract_recipe(final_fit)

preds<-final_fit$.workflow[[1]]$pre$actions$recipe$recipe$var_info %>%
 dplyr::filter(role=="predictor")

num_preds<-nrow(preds)


# Read in the out-of-sample data
# from data_prep.ml.R
# this contains ALL unclassifieds, however, only unclassifieds with mark_in==1 should get a prediction
oos_data<-readr::read_rds(file=here("data_folder","main","commercial",glue("BSB_unclassified_dataset{raw_oos_data_vintage_string}.Rds")))
#tack on original market_category and SPECIES_ITIS
oos_data<-oos_data %>%
  mutate(original_market_category=market_desc,
         SPECIES_ITIS="167687")


oos_predictable<-oos_data %>%
  filter(mark_in==1) 

oos_unpredictable<-oos_data %>%
  filter(mark_in!=1) 



# Read in calibrationmodel 
calibrate_iso<-readr::read_rds(file=here("results","ranger","final",glue("calibrate_iso_{modeltype}{finalfit_vintage}.Rds")))

```

## Predict and calibrate

Apply the final RF model to the data. Calibrate the probabilities

```{r Validation_predictions}
# Apply the Final fit to the out of sample dataset that I feel confident about 
oos_predictable <- augment(final_workflow,oos_predictable)


# Apply calibration apply to the out-of-sample dataset

oos_data_calibration_applied <-
  oos_predictable %>%
  cal_apply(calibrate_iso)

oos_data_calibration_applied <-oos_data_calibration_applied  %>%
  mutate(SEMESTER=as.numeric(as.character(semester))) %>%
  select(-semester)
```



## Construct predicted landings

Weight each observation by live pounds (NOT lndlb).  I've subsequently figured out better looking code to do this (predictions_heatmap), but this will work.

```{r extract_predictions}
### pull a few categories, compute a prob*lndlb variable, sum 


# pull the probability names
prob_names<-colnames(oos_data_calibration_applied) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)




# keep just a few columns
oos_data_calibration_applied<-oos_data_calibration_applied %>%
  select(c(year, stockarea, market_desc, original_market_category, SPECIES_ITIS, SEMESTER, weighting, livlb, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the live pounds to get predicted live pounds in each class. 
# This loop is a ugly.  There's only 5 categories, so it's probably better to just write it out
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    oos_data_calibration_applied[[paste0("pred_", l)]] <- oos_data_calibration_applied[[paste0(".pred_", l)]] * oos_data_calibration_applied$livlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}

# Sum up the predictions by year, market category, and stockarea. 
YRS_predictions<-oos_data_calibration_applied %>%
   group_by(year, stockarea,SEMESTER, original_market_category, SPECIES_ITIS) %>%
   mutate(transactions=1) %>%
   summarise(across(c(starts_with(c("pred_"))), \(x) sum(x, na.rm = TRUE))) %>%
  ungroup() %>%
    select(year,SEMESTER, stockarea, original_market_category, SPECIES_ITIS, starts_with("pred_"))

#reshape to a tidy long format
YRS_predictions<-  YRS_predictions %>%
      pivot_longer(
        cols = c(starts_with("pred_")),
        names_to = "Market.Comb",
        values_to = "live_lbs",
        names_prefix="pred_"
      ) %>%
  mutate(live_lbs=live_lbs/lbs_per_mt)%>%
  rename(live_metric_tons=live_lbs)

oos_unpredictable<-oos_unpredictable %>%
  mutate(SEMESTER=as.numeric(as.character(semester))) %>%
  select(-semester)

# get the unpredictables to the same format
oos_Uncs<-oos_unpredictable %>%
   group_by(year, stockarea,SEMESTER,original_market_category, SPECIES_ITIS ) %>%
   summarise(live_metric_tons=sum(livlb)/lbs_per_mt) %>%
   mutate(Market.Comb="Unclassified")

#bind together
YRS_predictions<-YRS_predictions %>%
  bind_rows(oos_Uncs)
  
  
# data quality check
# ensure I have 1 observation per year-region-SEMESTER
check1<-YRS_predictions %>%
  group_by(year,SEMESTER,stockarea, Market.Comb, original_market_category, SPECIES_ITIS)%>%
  summarise(count=n()) %>%
  ungroup()%>%
  summarise(max=max(count))
stopifnot(check1==1)



YRS_predictions<-YRS_predictions %>%
  mutate(Market.Comb=fct_relevel(Market.Comb,"Jumbo","Large","Medium","Small","Unclassified")) %>%
  group_by(year,SEMESTER,stockarea, Market.Comb, original_market_category, SPECIES_ITIS)%>%
  arrange(SPECIES_ITIS, year,SEMESTER,stockarea,original_market_category, Market.Comb)%>%
  ungroup() %>%
 mutate(YEAR = as.numeric(as.character(year)),
           STOCK_ABBREV = toupper(stockarea),
           MARKET_DESC = toupper(Market.Comb),
           MARKET_DESC = case_when (MARKET_DESC == 'MEDIUM' ~ "MEDIUM OR SELECT",
                                    MARKET_DESC != 'MEDIUM' ~ MARKET_DESC),
           MARKET_DESC_ORIG = toupper(original_market_category),
           LANDINGS_KG_CATEGORY_APPORTION = live_metric_tons*1000) %>% 
    select(YEAR,SEMESTER, STOCK_ABBREV,MARKET_DESC_ORIG, SPECIES_ITIS,MARKET_DESC,LANDINGS_KG_CATEGORY_APPORTION) 


write_rds(YRS_predictions, file=here("data_folder","predictions",glue("out_of_sample_predictions_YRS_{modeltype}{finalfit_vintage}.Rds")))


# Column Names
# This gets consumed by the LAA_calculation function, so don't change the variable names or things will break downstream.
#' \itemize{
#'   \item{YEAR}
#'   \item{SEMESTER}
#'   \item{STOCK_ABBREV} 
#'   \item{MARKET_DESC_ORIG}
#'   \item{SPECIES_ITIS}
#'   \item{MARKET_DESC}
#'   \item{LANDINGS_KG_CATEGORY_APPORTION}
#' }
#' 


```

## Ambitious predictions

I refrained from making some predictions about the unclassifieds, because they are similar to the data that I excluded from fitting the initial model. What happens if I try to predict class for those observations?  This repeats most of the previous, except it reads in ALL the oos_data, instead of just the mark_in==1

```{r ambitious_predictions}

# Apply the Final fit to the out of sample dataset 
ambitious_predictions <- augment(final_workflow,oos_data)

# apply Calibration

ambitious_predictions_cal_applied <-
  ambitious_predictions %>%
  cal_apply(calibrate_iso)

ambitious_predictions_cal_applied <-ambitious_predictions_cal_applied %>%
  mutate(SEMESTER=as.numeric(as.character(semester))) %>%
  select(-semester)


write_rds(ambitious_predictions_cal_applied, file=here("results","ranger",glue("out_of_sample_predictions_disaggregated_{modeltype}{finalfit_vintage}.Rds")))

# keep just a few columns
ambitious_predictions_cal_applied<-ambitious_predictions_cal_applied %>%
  select(c(year, stockarea, market_desc, SEMESTER, original_market_category, SPECIES_ITIS, weighting, livlb, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the landed pounds to get predicted pounds in each class. 
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    ambitious_predictions_cal_applied[[paste0("pred_", l)]] <- ambitious_predictions_cal_applied[[paste0(".pred_", l)]] * ambitious_predictions_cal_applied$livlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}

# Sum up the predictions by year, market category, and stockarea. 
ambitious_YRS_predictions<-ambitious_predictions_cal_applied %>%
   group_by(year, stockarea,SEMESTER,original_market_category, SPECIES_ITIS) %>%
   mutate(transactions=1) %>%
   summarise(across(c(starts_with(c("pred_"))), \(x) sum(x, na.rm = TRUE))) %>%
  ungroup() %>%
    select(year,SEMESTER, stockarea, original_market_category, SPECIES_ITIS,starts_with("pred_"))

ambitious_YRS_predictions<-  ambitious_YRS_predictions %>%
      pivot_longer(
        cols = c(starts_with("pred_")),
        names_to = "Market.Comb",
        values_to = "live_lbs",
        names_prefix="pred_"
      ) %>%
  mutate(live_lbs=live_lbs/lbs_per_mt)%>%
  rename(live_metric_tons=live_lbs)



# ensure I have 1 observation per year-region-SEMESTER
check1<-ambitious_YRS_predictions %>%
  group_by(year,SEMESTER,stockarea, original_market_category, SPECIES_ITIS, Market.Comb)%>%
  summarise(count=n()) %>%
  ungroup()%>%
  summarise(max=max(count))
stopifnot(check1==1)

ambitious_YRS_predictions<-ambitious_YRS_predictions %>%
  mutate(Market.Comb=fct_relevel(Market.Comb,"Jumbo","Large","Medium","Small")) %>%
  group_by(year,SEMESTER,stockarea, Market.Comb,original_market_category, SPECIES_ITIS)%>%
  arrange(SPECIES_ITIS, year,SEMESTER,stockarea,original_market_category, Market.Comb)%>%
  ungroup() %>%
 mutate(YEAR = as.numeric(as.character(year)),
           STOCK_ABBREV = toupper(stockarea),
           MARKET_DESC = toupper(Market.Comb),
           MARKET_DESC = case_when (MARKET_DESC == 'MEDIUM' ~ "MEDIUM OR SELECT",
                                    MARKET_DESC != 'MEDIUM' ~ MARKET_DESC),
           MARKET_DESC_ORIG = toupper(original_market_category),
           LANDINGS_KG_CATEGORY_APPORTION = live_metric_tons*1000) %>% 
    select(YEAR,SEMESTER, STOCK_ABBREV,MARKET_DESC_ORIG, SPECIES_ITIS,MARKET_DESC,LANDINGS_KG_CATEGORY_APPORTION) 

# Pad out the Ambitious classifications
padded_unc<-ambitious_YRS_predictions %>%
  group_by(YEAR, SEMESTER, STOCK_ABBREV, MARKET_DESC_ORIG, SPECIES_ITIS) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(MARKET_DESC="UNCLASSIFIED",
         LANDINGS_KG_CATEGORY_APPORTION=0)


padded<-rbind(ambitious_YRS_predictions, padded_unc)
padded<-padded %>%
  arrange(YEAR, SEMESTER, STOCK_ABBREV, MARKET_DESC_ORIG, SPECIES_ITIS, MARKET_DESC)
stopifnot(nrow(padded)==nrow(YRS_predictions))
#YRS -- year, region, SEMESTER

write_rds(padded, file=here("data_folder","predictions",glue("ambitious_out_of_sample_predictions_YRS_{modeltype}{finalfit_vintage}.Rds")))

# Column Names
# This gets consumed by the LAA_calculation function, so don't change the variable names or things will break downstream.
#' \itemize{
#'   \item{YEAR}
#'   \item{SEMESTER}
#'   \item{STOCK_ABBREV} 
#'   \item{MARKET_DESC_ORIG}
#'   \item{SPECIES_ITIS}
#'   \item{MARKET_DESC}
#'   \item{LANDINGS_KG_CATEGORY_APPORTION}
#' }
#' 



 ggplot(ambitious_YRS_predictions, aes(x=YEAR, y=LANDINGS_KG_CATEGORY_APPORTION, color=MARKET_DESC))+
    geom_line() +
  scale_x_continuous(name="year")  + 
  scale_y_continuous(name="Predicted live metric tons")  + 
  facet_grid(vars(STOCK_ABBREV), vars(SEMESTER)) 


```



```{r nicer_conservative_plot}

#YRS_predictions<-YRS_predictions %>%
#    mutate(stockarea = fct_rev(stockarea))


class_colours <- c(
  "JUMBO"  = "#1B6CA8",   # deep blue
  "LARGE"  = "#E05C2A",   # burnt orange
  "MEDIUM OR SELECT" = "#2E8B57",   # sea green
  "SMALL"  = "#7B3F9E" ,   # purple
  "UNCLASSIFIED" = "#404040" # dark gray
)

#scale landings_kg_category_apportion on the fly to mt
p_conserv_predictions <- ggplot(
    YRS_predictions ,
    aes(x = YEAR, y = LANDINGS_KG_CATEGORY_APPORTION/1000, fill = MARKET_DESC)
  ) +
  geom_col(
    position = position_stack(reverse = TRUE),
    width    = 0.8,       # leaves a visible gap between years
    colour   = NA         # suppress bar outlines
  ) +
  facet_grid(
    rows = vars(STOCK_ABBREV),
    cols = vars(SEMESTER)
  ) +
  scale_fill_manual(
    name   = "Market Category",
    values = class_colours,
    labels = c("Jumbo", "Large", "Medium", "Small","Unclassified")
  ) +
  scale_x_continuous(
    name   = "Year",
    breaks = scales::breaks_pretty(n = 5),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Predicted live weight (mt)",
    labels = scales::label_comma(accuracy = 1),
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_bw(base_size = 9) +
  theme(
    strip.background   = element_rect(fill = "grey92", colour = "grey40"),
    strip.text         = element_text(size = 8, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    axis.title         = element_text(size = 8),
    axis.text          = element_text(size = 7, colour = "grey20"),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    legend.position    = "bottom",
    legend.title       = element_text(size = 7, face = "bold"),
    legend.text        = element_text(size = 7),
    legend.key.width   = unit(0.4, "cm"),
    legend.key.height  = unit(0.3, "cm"),
    legend.margin      = margin(0, 0, 0, 0),
    legend.box.spacing = unit(4, "pt"),
    plot.margin        = margin(4, 6, 4, 4, "pt")
  )

ggsave(
  here("results", "ranger", "final",
       glue("YRS_predictions_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_conserv_predictions,
  width  = 174,
  height = 120,
  units  = "mm",
  device = cairo_pdf
)
p_conserv_predictions
```


```{r nicer_ambitious_plot}

#scale landings_kg_category_apportion on the fly to mt

p_predictions <- ggplot(
    ambitious_YRS_predictions,
    aes(x = YEAR, y = LANDINGS_KG_CATEGORY_APPORTION/1000, fill = MARKET_DESC)
  ) +
  geom_col(
    position = position_stack(reverse = TRUE),
    width    = 0.8,       # leaves a visible gap between years
    colour   = NA         # suppress bar outlines
  ) +
  facet_grid(
    rows = vars(STOCK_ABBREV),
    cols = vars(SEMESTER)
  ) +
  scale_fill_manual(
    name   = "Market grade",
    values = class_colours,
    labels = c("Jumbo", "Large", "Medium", "Small")
  ) +
  scale_x_continuous(
    name   = "Year",
    breaks = scales::breaks_pretty(n = 5),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Predicted live weight (mt)",
    labels = scales::label_comma(accuracy = 1),
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_bw(base_size = 9) +
  theme(
    strip.background   = element_rect(fill = "grey92", colour = "grey40"),
    strip.text         = element_text(size = 8, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    axis.title         = element_text(size = 8),
    axis.text          = element_text(size = 7, colour = "grey20"),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    legend.position    = "bottom",
    legend.title       = element_text(size = 7, face = "bold"),
    legend.text        = element_text(size = 7),
    legend.key.width   = unit(0.4, "cm"),
    legend.key.height  = unit(0.3, "cm"),
    legend.margin      = margin(0, 0, 0, 0),
    legend.box.spacing = unit(4, "pt"),
    plot.margin        = margin(4, 6, 4, 4, "pt")
  )

ggsave(
  here("results", "ranger", "final",
       glue("ambitious_YRS_predictions_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_predictions,
  width  = 174,
  height = 120,    # adjust in 5mm increments if strip labels or legend crowd
  units  = "mm",
  device = cairo_pdf
)

p_predictions
```






<!---
\newpage
--->
# References
<div id="refs"></div>





# Appendix{-}