---
title: "Construct Hard and Soft predictions for heatmaps"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: nocluster
    choices:
    - nocluster
    - TS_nocluster
---

# Summary and Housekeeping

<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	
Construct hard and soft predictive heatmaps.  Hard will go into the Appendix, soft goes into the manuscript.

* Outputs:
*   - Graphs of hard and soft predictive heatmaps.
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


**********************************************************************

 
 --->
```{r global_options, include=FALSE}

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/predictions_heatmap.Rmd")


# You've estimated a few difference models, this sets the 
if (modeltype=="standard"){
  data_pattern<-"data_split"
  tuning_pattern<-"BSB_ranger_tune"
  final_pattern<-"BSB_ranger_results"
} else if (modeltype=="nocluster"){
  data_pattern<-"nocluster_data_split"
  tuning_pattern<-"BSB_ranger_nocluster_tune"
  final_pattern<-"BSB_ranger_nocluster_results"
} else if (modeltype=="TS_nocluster"){
  data_pattern<-"nocluster_TS_data_split"
  tuning_pattern<-"BSB_ranger_TS_nocluster_tune"
  final_pattern<-"BSB_ranger_TS_nocluster_results"
}else {
  stop("Unknown modeltype")
}





#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)

# My "final fit" bombed out because I didn't have drive space.  
# finalfit_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{final_pattern}*Rds")))
# finalfit_vintage<-gsub(final_pattern,"",finalfit_vintage)
# finalfit_vintage<-gsub(".Rds","",finalfit_vintage)
# finalfit_vintage<-max(finalfit_vintage)

finalfit_vintage<-tuning_vintage

vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)



```

The modeltype is `r params$modeltype`, which is hopefully the same as `r modeltype`.

Data vintage is `r data_vintage_string`.

Tuning vintage is `r tuning_vintage`.

Final fit vintage is `r finalfit_vintage`.

```{r load_in_data and tuning_results, include=FALSE}
# this was created with data_prep_ml.Rmd
data_split<-readr::read_rds(file=here("results","ranger",glue("{data_pattern}{data_vintage_string}.Rds")))
#tune_res<-readr::read_rds(file=here("results","ranger",glue("{tuning_pattern}{tuning_vintage}.Rds")))

#train_data <- training(data_split)
test_data <- testing(data_split)
train_data <- training(data_split)

validation_data <- validation(data_split)

source(here("R_code","analysis","fit_random_forest","BSB.Classification.Recipe.R"))
search_type<-"Initial"


source(here("R_code","analysis","fit_random_forest","BSB.Workflow.Setup.R"))


```


# Load in the final model

read in the final model, extract the recipe and the predictors, extract the predictions

```{r load_final_model}

final_fit<-read_rds(file=here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))


# What is the workflow
final_workflow<-extract_workflow(final_fit)

# And the recipe
extract_recipe(final_fit)

preds<-final_fit$.workflow[[1]]$pre$actions$recipe$recipe$var_info %>%
 dplyr::filter(role=="predictor")

num_preds<-nrow(preds)

final_predictions<-final_fit %>%
  collect_predictions()

# pull the probability names
prob_names<-colnames(predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


# look at the fit metrics
final_fit_metrics<-final_fit %>%
  collect_metrics()
final_fit_metrics

# all.metrics %>%
#   dplyr::filter(.config==best_config) 

```

Log loss of the final fit is slightly better than that of the average of the individual folds. This is normal/reasonable, since it uses more data than each of the individual folds.  It's not so different that alarm bells go off.


# Testing and Validation predictions  

Use the Validation set to (1 assess how well the model fits) and 2. Calibrate if I have systematic under or over prediction.


```{r Validation_predictions}
# Load in the validation data

test_data <- augment(final_workflow,test_data)

```


```{r calibration_models}
# load in the calibration model

calibrate_iso<-readr::read_rds(file=here("results","ranger","final",glue("calibrate_iso_{modeltype}{finalfit_vintage}.Rds")))



```



```{r apply_calibration}
# apply to the test dataset

test_data_isocalib_applied <-
  test_data %>%
  cal_apply(calibrate_iso)

```

#  Heatmap

```{r heatmap}
# Make a pub ready heatmap
cm <- conf_mat(test_data_isocalib_applied, truth = market_desc, estimate = .pred_class)

cm_table <- as.data.frame(cm$table)
# Define desired order
class_levels <- c("Jumbo", "Large", "Medium", "Small")

cm_norm <- cm_table %>%
  group_by(Truth) %>%
  mutate(Prop = Freq / sum(Freq)) %>%
  ungroup() %>%
  mutate(
    Prediction = factor(Prediction, levels = class_levels),
    Truth      = factor(Truth,      levels = rev(class_levels))  # reversed for y-axis
  )

p_cm <- ggplot(cm_norm, aes(x = Prediction, y = Truth, fill = Prop)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(
    aes(label = sprintf("%.2f\n(n=%d)", Prop, Freq)),
    size   = 2.8,
    colour = "grey10"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#2166AC",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    name   = "Recall\n(row prop.)"
  ) +
  scale_x_discrete(position = "bottom") +
  labs(
    x = "Predicted class",
    y = "True class"
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    axis.text.x      = element_text(size = 7, colour = "grey20",
                                    angle = 45, hjust = 1),
    axis.text.y      = element_text(size = 7, colour = "grey20"),
    axis.title       = element_text(size = 8),
    legend.title     = element_text(size = 7),
    legend.text      = element_text(size = 7),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.3, "cm"),
    panel.grid       = element_blank(),
    panel.border     = element_rect(colour = "grey40", linewidth = 0.5),
    plot.margin      = margin(4, 4, 4, 4, "pt")
  )

ggsave(
  here("results", "ranger", "final",
       glue("confmat_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_cm,
  width  = 84,
  height = 84,
  units  = "mm",
  device = cairo_pdf
)
  
```


# Extra code

I'm leaving this extra code around to make my life easier when I want to make something different.

```{r extract_predictions}
### pull a few categories, compute a prob*lndlb variable, sum 


# keep just a few columns
test_data_calibration_applied<-test_data_calibration_applied %>%
  select(c( weighting, market_desc, lndlb, stockarea, year, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the landed pounds to get predicted pounds in each class. 
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    test_data_calibration_applied[[paste0("pred_", l)]] <- test_data_calibration_applied[[paste0(".pred_", l)]] * test_data_calibration_applied$lndlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}
# Sum up the predictions by year, market category, and stockarea. 
test_predictions<-test_data_calibration_applied %>%
   group_by(stockarea, year,market_desc) %>%
  select(-.pred_class) %>%
  mutate(transactions=1) %>%
  summarise(across(c(starts_with(c("pred_",".pred_")), "lndlb", "transactions"), \(x) sum(x, na.rm = TRUE))) %>%
   ungroup()
test_predictions$modeltype<-modeltype

test_predictions<-test_predictions %>%
  mutate(market_desc=fct_relevel(market_desc,"Jumbo","Large","Medium","Small")) 

write_rds(test_predictions, file=here("results","ranger",glue("aggregate_test_predictions_{modeltype}{finalfit_vintage}.Rds")))

#test_predictions<-test_predictions %>%
#  dplyr::filter(year %in% c(2021,2022,2023,2024))

```

## Probabalistic Predictions of the class of each transaction

Here are the predictions (transactions and mt) for the hold-out sample, using the probability calibrations.

```{r topline_validation}

aggregate_transactions<-test_predictions %>%
    summarise(across(c(starts_with(".pred_")), sum)) %>%
  pivot_longer(cols=starts_with(".pred_"), names_to="market_desc", names_prefix=".pred_",values_to="ObsPredicted")


aggregate_test_predictions<-test_predictions %>%
    summarise(across(c(starts_with("pred_")), sum)) %>%
  pivot_longer(cols=starts_with("pred_"), names_to="market_desc", names_prefix="pred_",values_to="predicted")


true<-test_predictions %>%
    group_by(market_desc) %>%
    summarise(lndlb=sum(lndlb),
              transactions=sum(transactions))

Testmkt_preds<-aggregate_transactions %>%
  left_join(true, by=join_by(market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  left_join(aggregate_test_predictions, by=join_by(market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted,
         predicted_transactions=ObsPredicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01),
         `Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )


transaction_predictions<-Testmkt_preds %>%
  select(market_desc,transactions, predicted_transactions, `Transaction error (%)`)


knitr::kable(transaction_predictions, caption='Calibrated Predictions on the 15% Validation Sample (transaction count).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```

```{r weighted_predictions}
weighted_predictions<-Testmkt_preds %>%
  select(market_desc,true_mt, predicted_mt, `mt error (%)`)

knitr::kable(weighted_predictions, caption='Calibrated Predictions on the 15% Validation Sample (mt).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```


Here are the predictions (transactions and mt) for the hold-out sample, without the probability calibrations.

```{r aggregate_test2_predictions}
# Subset data, construct predicted pounds, and aggregate


### pull a few categories, compute a prob*lndlb variable, sum 


# keep just a few columns
test_data_nocalibration_applied<-test_data%>%
  select(c( weighting, market_desc, lndlb, stockarea, year, any_of(prob_names), .pred_class))

# predicted pounds per transaction in the validation dataset.
# multiply the prediction by the landed pounds to get predicted pounds in each class. 
for (l in c("Jumbo", "Large", "Medium", "Small", "Unclassified")) {
  tryCatch({
    test_data_nocalibration_applied[[paste0("pred_", l)]] <- test_data_nocalibration_applied[[paste0(".pred_", l)]] * test_data_nocalibration_applied$lndlb
  }, error = function(e) {
    # R will just silently continue to the next iteration
  })
}
# Sum up the predictions by year, market category, and stockarea. 
nocal_predictions<-test_data_nocalibration_applied%>%
   group_by(stockarea, year,market_desc) %>%
  select(-.pred_class) %>%
  mutate(transactions=1) %>%
  summarise(across(c(starts_with(c("pred_",".pred_")), "lndlb", "transactions"), \(x) sum(x, na.rm = TRUE))) %>%
   ungroup()
nocal_predictions$modeltype<-modeltype

nocal_predictions<-nocal_predictions%>%
  mutate(market_desc=fct_relevel(market_desc,"Jumbo","Large","Medium","Small")) 

write_rds(nocal_predictions, file=here("results","ranger",glue("aggregate_nocal_predictions_{modeltype}{finalfit_vintage}.Rds")))

```







```{r no_cal_validations}

aggregate_nocal_transactions<-nocal_predictions%>%
    summarise(across(c(starts_with(".pred_")), sum)) %>%
  pivot_longer(cols=starts_with(".pred_"), names_to="market_desc", names_prefix=".pred_",values_to="ObsPredicted")


aggregate_nocal_weight_predictions<-nocal_predictions %>%
    summarise(across(c(starts_with("pred_")), sum)) %>%
  pivot_longer(cols=starts_with("pred_"), names_to="market_desc", names_prefix="pred_",values_to="predicted")


true<-test_data%>%
    mutate(transactions=1) %>%
    group_by(market_desc) %>%
    summarise(lndlb=sum(lndlb),
              transactions=sum(transactions))

nocalibrate_results<-aggregate_nocal_weight_predictions%>%
  left_join(true, by=join_by(market_desc==market_desc))

nocalibrate_results<-nocalibrate_results%>%
  left_join(aggregate_nocal_transactions, by=join_by(market_desc==market_desc))

nocalibrate_results<-nocalibrate_results%>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted,
         predicted_transactions=ObsPredicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01),
         `Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )
  

transactions_nocalibrate_predictions<-nocalibrate_results %>%
  select(market_desc,transactions, predicted_transactions, `Transaction error (%)`)


knitr::kable(transactions_nocalibrate_predictions, caption='UnCalibrated Predictions on the 15% Validation Sample (transactions).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```


```{r weighted_no_calibrate_predictions}
weights_nocalibrate_predictions<-nocalibrate_results %>%
  select(market_desc,true_mt, predicted_mt, `mt error (%)`)
knitr::kable(weights_nocalibrate_predictions, caption='UnCalibrated Predictions on the 15% Validation Sample (transactions).',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```









## Lets take a look at the validation predictions by stockarea



```{r Validation_Weight_Predictions_by_stockarea}

# Aggregate to the stockarea- market category

aggregate_test_predictions2<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(across(c(starts_with(".pred_"),lndlb), sum)) %>%
  arrange(stockarea,market_desc)

# true landings by stockarea and market_category
trueTest<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(lndlb=sum(lndlb))


# aggregate predictions, in pounds, by stockarea and market_category
Tp1<-test_predictions %>%
  group_by(stockarea) %>%
    summarise(across(c(starts_with("pred_")), sum))

Tp2<-Tp1  %>%
  pivot_longer(cols=!stockarea, names_to="market_desc", names_prefix="pred_",values_to="predicted")


Testmkt_preds<-Tp2 %>%
  left_join(trueTest, by=join_by(stockarea==stockarea, market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  mutate(predicted=predicted/lbs_per_mt,
         lndlb=lndlb/lbs_per_mt) %>%
  rename(true_mt=lndlb,
         predicted_mt=predicted)%>%
  mutate(`mt error (%)`=percent((predicted_mt-true_mt)/true_mt, accuracy=0.01) )


knitr::kable(Testmkt_preds, caption='Calibrated Predictions on the 15% Validation Sample (mt). Disaggregated to Stock area',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  

```




```{r Validation_Transaction_Predictions_by_stockarea}


# true landings by stockarea and market_category
trueTest<-test_predictions %>%
    group_by(stockarea,market_desc) %>%
    summarise(transactions=sum(transactions))


# aggregate predictions, in transactions, by stockarea and market_category
Tp1<-test_predictions %>%
  group_by(stockarea) %>%
    summarise(across(c(starts_with(".pred_")), sum))

Tp2<-Tp1  %>%
  pivot_longer(cols=!stockarea, names_to="market_desc", names_prefix=".pred_",values_to="predicted")


Testmkt_preds<-Tp2 %>%
  left_join(trueTest, by=join_by(stockarea==stockarea, market_desc==market_desc))

Testmkt_preds<-Testmkt_preds %>%
  rename(predicted_transactions=predicted) %>%
  mutate(`Transaction error (%)` = percent((predicted_transactions-transactions)/transactions, accuracy=0.01) )


knitr::kable(Testmkt_preds, caption='Calibrated Predictions on the 15% Validation Sample (mt). Disaggregated to Stock area',format.args = list(big.mark = ","), digits=0, align=c("l",rep('r',times=4)))  
```






<!---
\newpage
--->
# References
<div id="refs"></div>





# Appendix{-}