---
title: "Construct Hard and Soft predictions for heatmaps"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: nocluster
    choices:
    - nocluster
    - TS_nocluster
---

# Summary and Housekeeping

<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	
Construct hard and soft predictive heatmaps.  Hard will go into the Appendix, soft goes into the manuscript.

* Outputs:
*   - Graphs of hard and soft predictive heatmaps.
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


**********************************************************************

 
 --->
```{r global_options, include=FALSE}
search_type<-"Advanced"

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/predictions_heatmap.Rmd")

#handle model patterns
source(here("R_code","analysis","helpers","modeltype_patterns.R"))
source(here("R_code","analysis","helpers","predict_byhand.R"))




#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)

# Tuning, training, VI code guarantees that the finalfit and VI vintages are the same as the tuning vintages.  

finalfit_vintage<-tuning_vintage
VI_vintage<-tuning_vintage


final_fit_file_name<-glue("{final_pattern}{tuning_vintage}.Rds")
vi_file_name<-glue("{vi_pattern}{tuning_vintage}.Rds")
prepped_recipe_file_name<-glue("{prepped_recipe}{tuning_vintage}.Rds")


vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))


vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)

pred_vintage<-finalfit_vintage


```

The modeltype is `r params$modeltype`, which is hopefully the same as `r modeltype`.

Data vintage is `r data_vintage_string`.

Tuning vintage is `r tuning_vintage`.

Final fit vintage is `r finalfit_vintage`.

```{r load_in_data and tuning_results, include=FALSE}
# this was created with data_prep_ml.Rmd
data_split<-readr::read_rds(file=here("results","ranger",glue("{data_pattern}{data_vintage_string}.Rds")))
#tune_res<-readr::read_rds(file=here("results","ranger",glue("{tuning_pattern}{tuning_vintage}.Rds")))

test_data <- testing(data_split)
train_data <- training(data_split)

validation_data <- validation(data_split)
# First split was 80, 5, 15.  reasonable when I was going to unweight at all stages
# I think this isn't leaving enough data for the calibration 
# I'm going to re-stick the validation and test data together. Then I'm going to split it in half, so I get 
# 10% and 10%.

rebind<-rbind(test_data, validation_data)
# seed
set.seed(32560049)
ds2<-initial_split(rebind, prop=0.5 )

validation_data<-training(ds2)
test_data<-testing(ds2)



#source(here("R_code","analysis","fit_random_forest","BSB.Classification.Recipe.R"))

#source(here("R_code","analysis","fit_random_forest","BSB.Workflow.Setup.R"))


```


# Load in the final model

read in the final model, extract the recipe and the predictors, extract the predictions

```{r load_final_model}

prepped_recipe<-read_rds(file=here("results","ranger",prepped_recipe_file_name))

final_ranger_fit<-read_rds(file=here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))

# What the recipe

```

Log loss of the final fit is slightly better than that of the average of the individual folds. This is normal/reasonable, since it uses more data than each of the individual folds.  It's not so different that alarm bells go off.


# Confusion matrices and heatmaps

General strategy. Load in the final workflow and the holdout (test) data. load in the calibration model.
Predict and calibrate.

Make the heat maps (confusion matrices).




```{r Validation_predictions}
# Load in the validation data

################################################################################
# Generate class-probability predictions on the validation set via the custom
# predict_byhand() wrapper (applies prepped_recipe before calling ranger).
# The modal predicted class is extracted by finding the column with the highest
# probability for each row, then bound back alongside the raw probabilities
# and original validation data.

#predict
test_predictions<-predict_byhand(new_data=test_data,
                            prepped_recipe = prepped_recipe,
                            ranger_fitted_model  = final_ranger_fit)

class <- colnames(test_predictions)[max.col(test_predictions, ties.method = "first")]
class<-as_tibble(class) %>%
  rename(.pred_class=value) %>%
  mutate(.pred_class=str_remove(.pred_class,".pred_" ) # Removes all matches
)

class_levels <- c("Jumbo", "Large", "Medium", "Small")

class<-class %>%
  mutate(.pred_class=factor(.pred_class)) %>%
  mutate(.pred_class=fct_relevel(.pred_class,class_levels)
         )

# Bind cols
test_predictions<-bind_cols(test_predictions,test_data)
test_predictions<-bind_cols(class,test_predictions)

test_predictions<-test_predictions%>%
  mutate(weighting=hardhat::frequency_weights(lndlb)) 

# pull the probability names
prob_names<-colnames(test_predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


# all.metrics %>%
#   dplyr::filter(.config==best_config) 


```


```{r load_calibration}
# load in the calibration model

calibrate_weighted_multinom<-readr::read_rds(file=here("results","ranger","final",glue("calibrate_weighted_multinom_{modeltype}{finalfit_vintage}.Rds")))


```



```{r apply_calibration}
# apply to the test dataset

test_data_calib_applied <-
  test_predictions %>%
  cal_apply(calibrate_weighted_multinom)

```

#  Heatmaps

Make confusion matrix using the "conf_mat" function. 

Extract the data from that object and do some things to make it pub-ready.


```{r HeatmapHardPredictions}
# Make a pub ready heatmap based on the "hard" prediction (contained in .pred_class)
# These are "transaction-level, hard predictions" 
# the conf_mat object is not pub ready, so I've added some extra bits to make it look nice

# Pull the raw numbers and stick in a dataframe
cm <- conf_mat(test_data_calib_applied, truth = market_desc, estimate = .pred_class)
cm_table <- as.data.frame(cm$table)
# Define desired order
class_levels <- c("Jumbo", "Large", "Medium", "Small")

# Construct proportions
cm_norm <- cm_table %>%
  group_by(Truth) %>%
  mutate(Prop = Freq / sum(Freq)) %>%
  ungroup() %>%
  mutate(
    Prediction = factor(Prediction, levels = class_levels),
    Truth      = factor(Truth,      levels = rev(class_levels))  # reversed for y-axis
  )

# Plot
p_cm <- ggplot(cm_norm, aes(x = Prediction, y = Truth, fill = Prop)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(
    aes(label = paste0(sprintf("%.2f", Prop), "\n(",
                       formatC(Freq, format = "f", digits = 0, big.mark = ","),
                       ")")),
    size   = 2.8,
    colour = "grey10"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#2166AC",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    name   = "Recall\n(row prop.)"
  ) +
  scale_x_discrete(position = "bottom") +
  labs(
    x = "Predicted class",
    y = "True class"
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    axis.text.x      = element_text(size = 7, colour = "grey20",
                                    angle = 45, hjust = 1),
    axis.text.y      = element_text(size = 7, colour = "grey20"),
    axis.title       = element_text(size = 8),
    legend.title     = element_text(size = 7),
    legend.text      = element_text(size = 7),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.3, "cm"),
    panel.grid       = element_blank(),
    panel.border     = element_rect(colour = "grey40", linewidth = 0.5),
    plot.margin      = margin(4, 4, 4, 4, "pt")
  )
p_cm
ggsave(
  here("results", "ranger", "final",
       glue("confmat_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_cm,
  width  = 84,
  height = 84,
  units  = "mm",
  device = cairo_pdf
)
  
```


Same as previous, but make my own predictions based on the probability weights instead of using the default .class_pred.  


```{r HeatmapSoftPredictions}

# do "soft" predictions of Transactions by aggregating the predicted probabilities
soft_predictionUW<-test_data_calib_applied %>%
  group_by(market_desc) %>%
  summarise(Prediction_Jumbo=sum(.pred_Jumbo),
         Prediction_Large=sum(.pred_Large),
         Prediction_Medium=sum(.pred_Medium),
         Prediction_Small=sum(.pred_Small)
      )

# Reshape
soft_predictionUW <- soft_predictionUW %>%
  pivot_longer(
    cols = c(Prediction_Jumbo, Prediction_Large, Prediction_Medium, Prediction_Small),
    names_to = "Prediction",
    names_prefix = "Prediction_",
    values_to = "Freq"
  )
soft_predictionUW <-soft_predictionUW %>%
  rename(Truth=market_desc )

class_levels <- c("Jumbo", "Large", "Medium", "Small")

# compute proportions
sofc_cm <- soft_predictionUW %>%
  group_by(Truth) %>%
  mutate(Prop = Freq / sum(Freq)) %>%
  ungroup() %>%
  mutate(
    Prediction = factor(Prediction, levels = class_levels),
    Truth      = factor(Truth,      levels = rev(class_levels))  # reversed for y-axis
  )
# plot
p_softcm <- ggplot(sofc_cm, aes(x = Prediction, y = Truth, fill = Prop)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(
    aes(label = paste0(sprintf("%.2f", Prop), "\n(",
                       formatC(Freq, format = "f", digits = 0, big.mark = ","),
                       ")")),
    size   = 2.8,
    colour = "grey10"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#2166AC",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    name   = "Recall\n(row prop.)"
  ) +
  scale_x_discrete(position = "bottom") +
  labs(
    x = "Predicted class",
    y = "True class"
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    axis.text.x      = element_text(size = 7, colour = "grey20",
                                    angle = 45, hjust = 1),
    axis.text.y      = element_text(size = 7, colour = "grey20"),
    axis.title       = element_text(size = 8),
    legend.title     = element_text(size = 7),
    legend.text      = element_text(size = 7),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.3, "cm"),
    panel.grid       = element_blank(),
    panel.border     = element_rect(colour = "grey40", linewidth = 0.5),
    plot.margin      = margin(4, 4, 4, 4, "pt")
  )
p_softcm
ggsave(
  here("results", "ranger", "final",
       glue("softconfmat_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_softcm,
  width  = 84,
  height = 84,
  units  = "mm",
  device = cairo_pdf
)
```


```{r weighted_soft_conf_mat}
# do "soft" predictions of pounds by aggregating the predicted probabilities*lndlb

soft_predictionW<-test_data_calib_applied %>%
  group_by(market_desc) %>%
     summarise(Predlbs_Jumbo=sum(.pred_Jumbo*lndlb),
         Predlbs_Large=sum(.pred_Large*lndlb),
         Predlbs_Medium=sum(.pred_Medium*lndlb),
         Predlbs_Small=sum(.pred_Small*lndlb)
     )


# reshape
soft_predictionW <- soft_predictionW %>%
  pivot_longer(
    cols = c(Predlbs_Jumbo, Predlbs_Large, Predlbs_Medium, Predlbs_Small),
    names_to = "Prediction",
    names_prefix = "Predlbs_",
    values_to = "pounds"
  )


# rename columns, use metric tons
soft_predictionW <-soft_predictionW %>%
  rename(Truth=market_desc ) %>%
  mutate(metric_tons=pounds/lbs_per_mt)

class_levels <- c("Jumbo", "Large", "Medium", "Small")

soft_weighted <- soft_predictionW %>%
  group_by(Truth) %>%
  mutate(Prop = metric_tons / sum(metric_tons)) %>%
  ungroup() %>%
  mutate(
    Prediction = factor(Prediction, levels = class_levels),
    Truth      = factor(Truth,      levels = rev(class_levels))  # reversed for y-axis
  )

p_softcm <- ggplot(soft_weighted, aes(x = Prediction, y = Truth, fill = Prop)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(
    aes(label = paste0(sprintf("%.2f", Prop), "\n(",
                       formatC(metric_tons, format = "f", digits = 0, big.mark = ","),
                       ")")),
    size   = 2.8,
    colour = "grey10"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#2166AC",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    name   = "Recall\n(row prop.)"
  ) +
  scale_x_discrete(position = "bottom") +
  labs(
    x = "Predicted class",
    y = "True class"
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    axis.text.x      = element_text(size = 7, colour = "grey20",
                                    angle = 45, hjust = 1),
    axis.text.y      = element_text(size = 7, colour = "grey20"),
    axis.title       = element_text(size = 8),
    legend.title     = element_text(size = 7),
    legend.text      = element_text(size = 7),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.3, "cm"),
    panel.grid       = element_blank(),
    panel.border     = element_rect(colour = "grey40", linewidth = 0.5),
    plot.margin      = margin(4, 4, 4, 4, "pt")
  )
p_softcm
ggsave(
  here("results", "ranger", "final",
       glue("soft_weighted_confmat_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_softcm,
  width  = 84,
  height = 84,
  units  = "mm",
  device = cairo_pdf
)


```


















```{r save_the_predictions}
regional_predictions<-test_data_calib_applied %>%
  ungroup()%>%
  group_by(market_desc, year, stockarea,semester) %>%
  summarise(Prediction_Jumbo=sum(.pred_Jumbo),
         Prediction_Large=sum(.pred_Large),
         Prediction_Medium=sum(.pred_Medium),
         Prediction_Small=sum(.pred_Small),
         Predlbs_Jumbo=sum(.pred_Jumbo*lndlb),
         Predlbs_Large=sum(.pred_Large*lndlb),
         Predlbs_Medium=sum(.pred_Medium*lndlb),
         Predlbs_Small=sum(.pred_Small*lndlb),
         True_lbs=sum(lndlb),
         True_transactions=n(),
         .groups="drop_last"
  )


write_rds(regional_predictions, file=here("results","ranger",glue("soft_testing_predictions_{modeltype}{pred_vintage}.Rds")))

```

<!---
\newpage
--->
# References
<div id="refs"></div>





# Appendix{-}