---
title: "Construct Hard and Soft predictions for heatmaps"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: nocluster
    choices:
    - nocluster
    - TS_nocluster
---

# Summary and Housekeeping

<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	
Construct hard and soft predictive heatmaps.  Hard will go into the Appendix, soft goes into the manuscript.

* Outputs:
*   - Graphs of hard and soft predictive heatmaps.
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


**********************************************************************

 
 --->
```{r global_options, include=FALSE}

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/predictions_heatmap.Rmd")


# You've estimated a few difference models, this sets the 
if (modeltype=="standard"){
  data_pattern<-"data_split"
  tuning_pattern<-"BSB_ranger_tune"
  final_pattern<-"BSB_ranger_results"
} else if (modeltype=="nocluster"){
  data_pattern<-"nocluster_data_split"
  tuning_pattern<-"BSB_ranger_nocluster_tune"
  final_pattern<-"BSB_ranger_nocluster_results"
} else if (modeltype=="TS_nocluster"){
  data_pattern<-"nocluster_TS_data_split"
  tuning_pattern<-"BSB_ranger_TS_nocluster_tune"
  final_pattern<-"BSB_ranger_TS_nocluster_results"
}else {
  stop("Unknown modeltype")
}





#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)

# My "final fit" bombed out because I didn't have drive space.  
# finalfit_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{final_pattern}*Rds")))
# finalfit_vintage<-gsub(final_pattern,"",finalfit_vintage)
# finalfit_vintage<-gsub(".Rds","",finalfit_vintage)
# finalfit_vintage<-max(finalfit_vintage)

finalfit_vintage<-tuning_vintage

vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)

pred_vintage<-"2026-02-11"


```

The modeltype is `r params$modeltype`, which is hopefully the same as `r modeltype`.

Data vintage is `r data_vintage_string`.

Tuning vintage is `r tuning_vintage`.

Final fit vintage is `r finalfit_vintage`.

```{r load_in_data and tuning_results, include=FALSE}
# this was created with data_prep_ml.Rmd
data_split<-readr::read_rds(file=here("results","ranger",glue("{data_pattern}{data_vintage_string}.Rds")))
#tune_res<-readr::read_rds(file=here("results","ranger",glue("{tuning_pattern}{tuning_vintage}.Rds")))

#train_data <- training(data_split)
test_data <- testing(data_split)
train_data <- training(data_split)

validation_data <- validation(data_split)

source(here("R_code","analysis","fit_random_forest","BSB.Classification.Recipe.R"))
search_type<-"Initial"


source(here("R_code","analysis","fit_random_forest","BSB.Workflow.Setup.R"))


```


# Load in the final model

read in the final model, extract the recipe and the predictors, extract the predictions

```{r load_final_model}

final_fit<-read_rds(file=here("results","ranger",glue("{final_pattern}{tuning_vintage}.Rds")))


# What is the workflow
final_workflow<-extract_workflow(final_fit)

# And the recipe
extract_recipe(final_fit)

preds<-final_fit$.workflow[[1]]$pre$actions$recipe$recipe$var_info %>%
 dplyr::filter(role=="predictor")

num_preds<-nrow(preds)

final_predictions<-final_fit %>%
  collect_predictions()

# pull the probability names
prob_names<-colnames(predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


# look at the fit metrics
final_fit_metrics<-final_fit %>%
  collect_metrics()
final_fit_metrics

# all.metrics %>%
#   dplyr::filter(.config==best_config) 

```

Log loss of the final fit is slightly better than that of the average of the individual folds. This is normal/reasonable, since it uses more data than each of the individual folds.  It's not so different that alarm bells go off.


# Confusion matrices and heatmaps

General strategy. Load in the final workflow and the holdout (test) data. load in the calibration model.
Predict and calibrate.

Make the heat maps (confusion matrices).




```{r Validation_predictions}
# Load in the validation data

test_data <- augment(final_workflow,test_data)

```


```{r calibration_models}
# load in the calibration model

calibrate_iso<-readr::read_rds(file=here("results","ranger","final",glue("calibrate_iso_{modeltype}{finalfit_vintage}.Rds")))



```



```{r apply_calibration}
# apply to the test dataset

test_data_isocalib_applied <-
  test_data %>%
  cal_apply(calibrate_iso)

```

#  Heatmaps

Make confusion matrix using the "conf_mat" function. 

Extract the data from that object and do some things to make it pub-ready.

```{r heatmap}
# Make a pub ready heatmap
cm <- conf_mat(test_data_isocalib_applied, truth = market_desc, estimate = .pred_class)

cm_table <- as.data.frame(cm$table)
# Define desired order
class_levels <- c("Jumbo", "Large", "Medium", "Small")

cm_norm <- cm_table %>%
  group_by(Truth) %>%
  mutate(Prop = Freq / sum(Freq)) %>%
  ungroup() %>%
  mutate(
    Prediction = factor(Prediction, levels = class_levels),
    Truth      = factor(Truth,      levels = rev(class_levels))  # reversed for y-axis
  )

p_cm <- ggplot(cm_norm, aes(x = Prediction, y = Truth, fill = Prop)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(
    aes(label = paste0(sprintf("%.2f", Prop), "\n(",
                       formatC(Freq, format = "f", digits = 0, big.mark = ","),
                       ")")),
    size   = 2.8,
    colour = "grey10"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#2166AC",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    name   = "Recall\n(row prop.)"
  ) +
  scale_x_discrete(position = "bottom") +
  labs(
    x = "Predicted class",
    y = "True class"
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    axis.text.x      = element_text(size = 7, colour = "grey20",
                                    angle = 45, hjust = 1),
    axis.text.y      = element_text(size = 7, colour = "grey20"),
    axis.title       = element_text(size = 8),
    legend.title     = element_text(size = 7),
    legend.text      = element_text(size = 7),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.3, "cm"),
    panel.grid       = element_blank(),
    panel.border     = element_rect(colour = "grey40", linewidth = 0.5),
    plot.margin      = margin(4, 4, 4, 4, "pt")
  )
p_cm
ggsave(
  here("results", "ranger", "final",
       glue("confmat_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_cm,
  width  = 84,
  height = 84,
  units  = "mm",
  device = cairo_pdf
)
  
```

Same as previous, but make my own predictions based on the probability weights instead of using the default .class_pred.  


```{r soft_prediction_conf_matrix}
soft_predictions<-test_data_isocalib_applied %>%
  group_by(market_desc) %>%
  summarise(Prediction_Jumbo=sum(.pred_Jumbo),
         Prediction_Large=sum(.pred_Large),
         Prediction_Medium=sum(.pred_Medium),
         Prediction_Small=sum(.pred_Small),
         Predlbs_Jumbo=sum(.pred_Jumbo*lndlb),
         Predlbs_Large=sum(.pred_Large*lndlb),
         Predlbs_Medium=sum(.pred_Medium*lndlb),
         Predlbs_Small=sum(.pred_Small*lndlb)
         )
soft_predictionUW <- soft_predictions %>%
  select(c(market_desc, starts_with("Prediction"))) %>%
  pivot_longer(
    cols = c(Prediction_Jumbo, Prediction_Large, Prediction_Medium, Prediction_Small),
    names_to = "Prediction",
    names_prefix = "Prediction_",
    values_to = "Freq"
  )
soft_predictionUW <-soft_predictionUW %>%
  rename(Truth=market_desc )

class_levels <- c("Jumbo", "Large", "Medium", "Small")

sofc_cm <- soft_predictionUW %>%
  group_by(Truth) %>%
  mutate(Prop = Freq / sum(Freq)) %>%
  ungroup() %>%
  mutate(
    Prediction = factor(Prediction, levels = class_levels),
    Truth      = factor(Truth,      levels = rev(class_levels))  # reversed for y-axis
  )

p_softcm <- ggplot(sofc_cm, aes(x = Prediction, y = Truth, fill = Prop)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(
    aes(label = paste0(sprintf("%.2f", Prop), "\n(",
                       formatC(Freq, format = "f", digits = 0, big.mark = ","),
                       ")")),
    size   = 2.8,
    colour = "grey10"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#2166AC",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    name   = "Recall\n(row prop.)"
  ) +
  scale_x_discrete(position = "bottom") +
  labs(
    x = "Predicted class",
    y = "True class"
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    axis.text.x      = element_text(size = 7, colour = "grey20",
                                    angle = 45, hjust = 1),
    axis.text.y      = element_text(size = 7, colour = "grey20"),
    axis.title       = element_text(size = 8),
    legend.title     = element_text(size = 7),
    legend.text      = element_text(size = 7),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.3, "cm"),
    panel.grid       = element_blank(),
    panel.border     = element_rect(colour = "grey40", linewidth = 0.5),
    plot.margin      = margin(4, 4, 4, 4, "pt")
  )
p_softcm
ggsave(
  here("results", "ranger", "final",
       glue("softconfmat_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_softcm,
  width  = 84,
  height = 84,
  units  = "mm",
  device = cairo_pdf
)
```


```{r weighted_soft_conf_mat}

soft_predictionW <- soft_predictions %>%
  select(c(market_desc, starts_with("Predlbs_"))) %>%
  pivot_longer(
    cols = c(Predlbs_Jumbo, Predlbs_Large, Predlbs_Medium, Predlbs_Small),
    names_to = "Prediction",
    names_prefix = "Predlbs_",
    values_to = "pounds"
  )



soft_predictionW <-soft_predictionW %>%
  rename(Truth=market_desc ) %>%
  mutate(metric_tons=pounds/lbs_per_mt)

class_levels <- c("Jumbo", "Large", "Medium", "Small")

soft_weighted <- soft_predictionW %>%
  group_by(Truth) %>%
  mutate(Prop = metric_tons / sum(metric_tons)) %>%
  ungroup() %>%
  mutate(
    Prediction = factor(Prediction, levels = class_levels),
    Truth      = factor(Truth,      levels = rev(class_levels))  # reversed for y-axis
  )

p_softcm <- ggplot(soft_weighted, aes(x = Prediction, y = Truth, fill = Prop)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(
    aes(label = paste0(sprintf("%.2f", Prop), "\n(",
                       formatC(metric_tons, format = "f", digits = 0, big.mark = ","),
                       ")")),
    size   = 2.8,
    colour = "grey10"
  ) +
  scale_fill_gradient(
    low    = "white",
    high   = "#2166AC",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    name   = "Recall\n(row prop.)"
  ) +
  scale_x_discrete(position = "bottom") +
  labs(
    x = "Predicted class",
    y = "True class"
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    axis.text.x      = element_text(size = 7, colour = "grey20",
                                    angle = 45, hjust = 1),
    axis.text.y      = element_text(size = 7, colour = "grey20"),
    axis.title       = element_text(size = 8),
    legend.title     = element_text(size = 7),
    legend.text      = element_text(size = 7),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.3, "cm"),
    panel.grid       = element_blank(),
    panel.border     = element_rect(colour = "grey40", linewidth = 0.5),
    plot.margin      = margin(4, 4, 4, 4, "pt")
  )
p_softcm
ggsave(
  here("results", "ranger", "final",
       glue("soft_weighted_confmat_{modeltype}{finalfit_vintage}.pdf")),
  plot   = p_softcm,
  width  = 84,
  height = 84,
  units  = "mm",
  device = cairo_pdf
)


```


















```{r save_the_predictions}
regional_predictions<-test_data_isocalib_applied %>%
  ungroup()%>%
  group_by(market_desc, year, stockarea,semester) %>%
  summarise(Prediction_Jumbo=sum(.pred_Jumbo),
         Prediction_Large=sum(.pred_Large),
         Prediction_Medium=sum(.pred_Medium),
         Prediction_Small=sum(.pred_Small),
         Predlbs_Jumbo=sum(.pred_Jumbo*lndlb),
         Predlbs_Large=sum(.pred_Large*lndlb),
         Predlbs_Medium=sum(.pred_Medium*lndlb),
         Predlbs_Small=sum(.pred_Small*lndlb),
         True_lbs=sum(lndlb),
         True_transactions=n()
         )%>%
  ungroup()


write_rds(regional_predictions, file=here("results","ranger",glue("soft_testing_predictions_{modeltype}{pred_vintage}.Rds")))

```

```{r save_the_predictions}
# create a fake dataset with the structure of the out of sample predictions for Emily.
mockup <-regional_predictions %>%
  select(year,semester, stockarea, starts_with("Predlbs"))

  
  
mock2<-  mockup %>%
      pivot_longer(
        cols = c(starts_with("Predlbs")),
        names_to = "Market.Comb",
        values_to = "lbs",
        names_prefix="Predlbs_"
      ) %>%
  mutate(lbs=lbs/lbs_per_mt)%>%
  rename(metric_tons=lbs)

write_rds(mock2, file=here("results","ranger",glue("mockup_predictions_{modeltype}{pred_vintage}.Rds")))


  
  
```
<!---
\newpage
--->
# References
<div id="refs"></div>





# Appendix{-}