---
title: "Black Sea Bass: tuning Diagnostics"
author: "Min-Yang Lee"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    keep_md: yes
    fig_caption: yes
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
header-includes: \usepackage{setspace}\doublespacing
urlcolor: blue
editor_options:
  chunk_output_type: console
fontsize: 12pt
params:
  modeltype:
    value: nocluster
    choices:
    - nocluster
    - noclusterTEST
---

# Summary and Housekeeping


 My "final fit" bombed out because I didn't have drive space.  This is code to
 
1. Read in the tune result and eyeball them.
2. Pick the best set of tuning parameters.
3. Fit a final model. 
4. make some figures that characterize the hyperparamter tuning process.



<!---- 
 The global_options chunk loads libraries, sets options, figures out if you're on a desktop or server, sets years, and sets graphing options
 **********************************************************************
* Purpose: 	Read in outputs of machine learning models 
* Inputs: These three can be adjusted by changing the "modeltype"
*   - Data file
*   - tuning results file
*   - final results file


* Outputs:
*   - Graphs of tuning quality and final model fit
*   - Aggregate predictions for the testing (validation) dataset and the entire dataset (train+test)


**********************************************************************

 
 --->
```{r global_options, include=FALSE}
search_type<-"Advanced"

modeltype<-params$modeltype


library("here")

# load tidyverse and related
library("tidyverse")
library("scales")
library("ggrepel")
# load tidyverse and related
library("tidymodels")

# load machine learning and estimation tools
library("nnet")
library("ranger")
library("partykit")
library("bonsai")
library("vip")
library("probably")
library("discrim")
library("betacal")

# load utilities
library("knitr")
library("kableExtra")
library("viridis")
library("glue")

#3d plots
library("htmlwidgets")
library("plotly")

library("conflicted")


#deal with conflicts
conflicts_prefer(dplyr::filter())
conflicts_prefer(dplyr::lag())
conflicts_prefer(purrr::discard())
conflicts_prefer(dplyr::group_rows())
conflicts_prefer(yardstick::spec())
conflicts_prefer(recipes::fixed())
conflicts_prefer(recipes::step())
conflicts_prefer(viridis::viridis_pal())

# Determine what platform the code is running on and set the number of threads for ranger
platform <- Sys.info()['sysname']
# check the name of the effective_user
if(platform == 'Linux'){
  if (grep("PREEMPT_DYNAMIC",Sys.info()['version'])==1){
      runClass<-'DynamicContainer'
    } else{ 
      runClass <- 'Container'
    }
  }
 

if(platform == 'Windows'){
  runClass<-'Windows'
}

if (runClass %in% c('Local', 'Windows')){
  my.parallel.threads<-parallel::detectCores() -2 
} else if (runClass %in% c('Container','DynamicContainer')){ 
					  
												
  my.parallel.threads<-2
  
}
my.ranger.threads<-7

here::i_am("writing/final_fit.Rmd")

source(here("R_code","analysis","helpers","modeltype_patterns.R"))
source(here("R_code","analysis","helpers","predict_byhand.R"))





#############################################################################
#knitr options

knitr::opts_chunk$set(echo=TRUE, warning = FALSE, error = FALSE, message = FALSE, comment = FALSE, cache = FALSE, progress = TRUE, verbose = FALSE, dev="png",
											dpi = 600)
options(tinytex.verbose = TRUE)
# options(knitr.table.format = "latex")
options(scipen=999)

lbs_per_mt<-2204.62
#############################################################################
my_images<-here("images")
descriptive_images<-here("images","descriptive")
exploratory_images<-here("images","exploratory")


data_vintage_string<-list.files(here("results","ranger"), pattern=glob2rx(glue("{data_pattern}*Rds")))
data_vintage_string<-gsub(data_pattern,"",data_vintage_string)
data_vintage_string<-gsub(".Rds","",data_vintage_string)
data_vintage_string<-max(data_vintage_string)


tuning_vintage<-list.files(here("results","ranger"), pattern=glob2rx(glue("{tuning_pattern}*Rds")))
tuning_vintage<-gsub(tuning_pattern,"",tuning_vintage)
tuning_vintage<-gsub(".Rds","",tuning_vintage)
tuning_vintage<-max(tuning_vintage)
# Tuning, training, VI code guarantees that the finalfit and VI vintages are the same as the tuning vintages.  

finalfit_vintage<-tuning_vintage
VI_vintage<-tuning_vintage

best_param_file_name<-glue("{best_param_pattern}{tuning_vintage}.Rds")

final_fit_file_name<-glue("{final_pattern}{tuning_vintage}.Rds")
vi_file_name<-glue("{vi_pattern}{tuning_vintage}.Rds")
prepped_recipe_file_name<-glue("{prepped_recipe}{tuning_vintage}.Rds")

vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_estimation_dataset*Rds"))

raw_oos_data_vintage_string<-list.files(here("data_folder","main","commercial"), pattern=glob2rx("BSB_unclassified_dataset*Rds"))
raw_oos_data_vintage_string<-gsub("BSB_unclassified_dataset","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-gsub(".Rds","",raw_oos_data_vintage_string)
raw_oos_data_vintage_string<-max(raw_oos_data_vintage_string)



```

The modeltype is `r params$modeltype`, which is hopefully the same as `r modeltype`.

Data vintage is `r data_vintage_string`.

Tuning vintage is `r tuning_vintage`.

Final fit vintage is `r finalfit_vintage`.

```{r load_in_tuning_results, include=FALSE}
tune_res<-readr::read_rds(file=here("results","ranger",glue("{tuning_pattern}{tuning_vintage}.Rds")))
metrics_by_fold<-read_rds(file=here("results","ranger", glue("tuning_metrics_by_fold{tuning_vintage}.Rds")))
best_params<-read_rds(file=here("results","ranger", best_param_file_name))

```


# Tuning

How well did the tuning work?

```{r check_convergence}

# Select the best params from tune_res

best_tree <- tune_res %>%
  select_best(metric = "brier_class")

best_config<-best_tree %>%
  pull(.config)

best_config


# I chose to not use the best params from tune res. The 2nd best model was very close
# in terms of brier_class, had some good 'speed' attributes. having a higher min_n also
# makes sense from a theory perspective with the uncounted data
tm<-metrics_by_fold  %>%
  filter(.metric == "brier_class") %>%
  group_by(mtry, min_n, .config) %>%
  summarise(mt=mean(.estimate), .groups="drop_last")%>%
  arrange(mt)

best_config<-tm[1,] %>%
  select(-mt)
best_config

trained_config<-tm[2,] %>%
  select(-mt)


trained_config

tm[1:15,]

```

Are there any warnings from the tuning?

```{r warnings}
extracts<-tune_res %>%
  collect_notes()
extracts
not_good<-unique(extracts$id)
not_good
```

# Some in-sample results from Tuning

The model tuning and training pipeline is different now. I've tuned the model on the "expanded data."  This means that the subsequent ROC and PR curves are "weighted" in the sense that there are repeated observations of data.

```{r roc}
best_predictions<-tune_res %>%
  collect_predictions(parameters=best_config)

trained_predictions<-tune_res %>%
  collect_predictions(parameters=trained_config)

#rm(tune_res)

# pull the probability names
prob_names<-colnames(trained_predictions) 
prob_names<-grep("^\\.pred_", prob_names, value=TRUE)
prob_names<-grep("^\\.pred_class", prob_names, value=TRUE, invert=TRUE)


roc_data<-trained_predictions %>%
  group_by(id) %>%
  roc_curve(truth=market_desc, 
            any_of(prob_names)

# --- 3. Colour palette — four classes ---
class_colours <- c(
  "Jumbo"  = "#1B6CA8",   # deep blue
  "Large"  = "#E05C2A",   # burnt orange
  "Medium" = "#2E8B57",   # sea green
  "Small"  = "#7B3F9E"    # purple
)

roc_mean <- roc_data %>%
  group_by(.level, specificity) %>%
  summarise(sensitivity = mean(sensitivity), .groups = "drop")


# --- 4. Build the figures ---
p_facet <- ggplot(roc_data,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level, group = interaction(.level, id)),
            linewidth = 0.2, alpha = 0.40, show.legend = FALSE) +
#  geom_path(data = roc_mean,
 #        aes(colour = "grey92"),
  #      linewidth = 0.25, alpha = 1.0, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
# --- 5. Save at ICES JMS specifications ---
ggsave(here("results","ranger","tune",glue("roc_tune_selected_{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facet,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)



roc_data_best_config<-best_predictions %>%
  group_by(id) %>%
  roc_curve(truth=market_desc, 
            any_of(prob_names)) 


roc_mean_best_config <- roc_data_best_config %>%
  group_by(.level, specificity) %>%
  summarise(sensitivity = mean(sensitivity), .groups = "drop")


# --- 3. Colour palette — four classes ---
class_colours <- c(
  "Jumbo"  = "#1B6CA8",   # deep blue
  "Large"  = "#E05C2A",   # burnt orange
  "Medium" = "#2E8B57",   # sea green
  "Small"  = "#7B3F9E"    # purple
)

# --- 4. Build the figures ---
p_facet <- ggplot(roc_data_best_config,
                  aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_path(aes(colour = .level, group = interaction(.level, id)),
            linewidth = 0.2, alpha = 0.40, show.legend = FALSE) +
  facet_wrap(~ .level, ncol = 4) +
  scale_colour_manual(values = class_colours) +
  scale_x_continuous(
    name   = "1 \u2212 Specificity (False Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  scale_y_continuous(
    name   = "Sensitivity (True Positive Rate)",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.25),
    expand = expansion(mult = 0.01)
  ) +
  coord_equal() +
  theme_bw(base_size = 9) +
  theme(
    strip.background = element_rect(fill = "grey92", colour = "grey40"),
    strip.text       = element_text(size = 8, face = "bold"),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    axis.title       = element_text(size = 8),
    axis.text        = element_text(size = 7, colour = "grey20"),
    plot.margin      = margin(4, 6, 4, 4, "pt")
  )
# --- 5. Save at ICES JMS specifications ---
ggsave(here("results","ranger","tune",glue("roc_tune_nonselected_{modeltype}{finalfit_vintage}.pdf")),
       plot   = p_facet,
       width  = 174,   # mm — double-column, suits 4 facets in a row
       height = 65,
       units  = "mm",
       device = cairo_pdf)








pr_plot<-trained_predictions %>%
  group_by(id) %>%
  pr_curve(truth=market_desc, any_of(prob_names))  %>%
 ggplot(aes(x=recall,y=precision, colour=id)) +
   facet_wrap(~.level) + 
  geom_line(show.legend = TRUE) 

ggsave(here("results","ranger","tune",glue("pr_curve_tune_{modeltype}{tuning_vintage}.png")), plot=pr_plot)

rm(pr_plot)
  rm(predictions)
gc()
```
